### 2023-01-17 1st_Manuscript/SupMaterial.Rmd
January 2023 updates

---
title: "Supplementary material for 'The complex network of trophic interactions in a subAntarctic Marine Protected Area'"
author: "Tomás I. Marina, Irene R. Schloss and Luciana Riccialdelli"
output: pdf_document
bibliography: "AMPBancoBurdwood.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## Food web construction

The following figure and tables summarize the data we used to build the network of trophic (predator-prey) interactions for the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem (MPAN-BB).
Figure 1 resumes the type and number of references reviewed. Table 1 is the complete list of trophic interactions and the references that confirm each of them. Table 2 is the species list with details on trophospecies cases (aggregated taxa).

```{r ref-plot, echo=FALSE, fig.cap="Summary of the references used to build the network of trophic (predator-prey) interactions for the MPAN-BB ecosystem."}

# Load packages
library(tidyverse)
library(ggplot2)

# Load data
load("../data/cleaned-data_jan23.rda")

# Discard low resolution interactions
int_good_res <- int_raw %>% 
  filter(!str_detect(Prey, "\\*$")) %>% 
  filter(!str_detect(Predator, "\\*$")) %>% 
  dplyr::select(Prey, Predator, PreyGroup, PredGroup, Reference) %>% 
  distinct(Prey, Predator, .keep_all = TRUE)

# Figure 1. Plot resuming references
biblio <- data.frame(unique(int_good_res$Reference))
colnames(biblio) <- "Reference"
biblio <- biblio %>% 
  mutate(Type = case_when(str_detect(Reference, "pers.com.") ~ "PersComm",
                          str_detect(Reference, "etal") ~ "Article",
                          str_detect(Reference, "&") ~ "Article",
                          str_detect(Reference, "Encyclopedia of Life") ~ "DataBase",
                          str_detect(Reference, "GloBI") ~ "DataBase",
                          str_detect(Reference, "GATEWAy") ~ "DataBase",
                          str_detect(Reference, "PhDTesis") ~ "Thesis",
                          TRUE ~ "Article"))
ggplot(biblio) +
  geom_bar(aes(x = Type, fill = Type)) +
  theme_bw() +
  labs(x = "Reference", y = "Count") +
  annotate("text", x = Inf, y = Inf, label = paste("Total = ", nrow(biblio), sep = ""), vjust=1, hjust=1)

```

\newpage

```{r echo=FALSE, results='asis'}

# Load packages
library(knitr)
library(dplyr)
library(kableExtra)
library(pander)

# Load data
load("../data/cleaned-data_jan23.rda")

# Discard low resolution interactions
int_good_res <- int_raw %>% 
  filter(!str_detect(Prey, "\\*$")) %>% 
  filter(!str_detect(Predator, "\\*$")) %>% 
  dplyr::select(Prey, Predator, Reference, Reference_link) %>% 
  distinct(Prey, Predator, .keep_all = TRUE)

# Table 1. List of interactions
knitr::kable(int_good_res %>% arrange(., (Predator)), booktabs = TRUE, format = "latex", longtable = T, col.names = c("Prey","Predator","Reference", "Link"), caption = "List of predator-prey (trophic) interactions used to build the network for the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem (MPAN-BB).") %>% 
  column_spec(1:3, width="2cm") %>% 
  column_spec(4, width="14cm") %>% 
  kable_styling(latex_options = c("hold_position","repeat_header"), font_size = 4) %>% 
  landscape()

# pandoc.table(int_good_res %>% arrange(., (Predator)) %>% mutate(Reference_link = paste('\\tiny', Reference_link)) %>% rename("Link"="Reference_link"), style="multiline", split.tables = Inf, justify = c("left","left","left","right"),
#              caption = "List of predator-prey (trophic) interactions used to build the network for the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem (MPAN-BB).")

```

\newpage

```{r echo=FALSE, results='asis'}

# Load packages
library(knitr)
library(dplyr)
library(kableExtra)

# Load data
load("../data/cleaned-data_jan23.rda")
load("../data/foodweb-data_jan23.rda")

# Tidy data
df_g <- igraph::as_data_frame(g, 'both')
sp_list <- df_g[["vertices"]] %>% 
  rename("TrophicSpecies" = "name") %>% 
  left_join(sp_raw) %>% 
  dplyr::select(TrophicSpecies, FunctionalGroup, Reference_pres, Reference_pres_link, Observation) %>% 
  rename("Species" = "TrophicSpecies", "Reference" = "Reference_pres") %>% 
  mutate(Observation = case_when(Observation %in% c("Registrado Psolus_sp; Martinez sugiere presencia de P. segregatus", "Registrado Psolus_sp; Martinez sugiere presencia de P. patagonicus", "Registrado Psolus_sp; Martinez sugiere presencia de P. antarcticus", "Lista de sp no publicada provista por Díaz de Astarloa", "Info from Sup Info (Teso_etal_2019)", "AMPNBB-II", "Abundance data available ?", "Esquema trófico", "Depredadores tope", "Copepoda", "Benthic. Not recorded", "Cirripedia") ~ "", 
                                 Species == "Campylaspis_spp" ~ "Aggregated: Campylaspis_frigida y Campylaspis_bacescui",
                                 Species == "Ophiactis_sp" ~ "Aggregated: Ophiactis_asperula y Ophiactis_spp",
                                 Species == "Brachiopoda_spp" ~ "Aggregated: Magellania_venosa, Liothyrella_uva y Terebratella_dorsata",
                                 TRUE ~ Observation),
         Reference = case_when(Species == "Bacteria" ~ "Guinder_etal_2020",
                               Species == "Necromass" ~ "Riccialdelli_etal_2017", 
                               Species == "Syllidae" ~ "Schejter_etal_2016; Bremec_etal_2019",
                               Species == "Sabellidae" ~ "Schejter_etal_2016; Bremec_etal_2019; Spinelli_etal_2020",
                               Species == "Ophiactis_sp" ~ "Schejter_etal_2016; 2020a; Brogger_pers.com.",
                               Species == "Campylaspis_spp" ~ "Doti_etal_2020",
                               Species == "Brachiopoda_spp" ~ "Schejter_etal_2016", TRUE ~ Reference),
         FunctionalGroup = case_when(Species == "Campylaspis_spp" ~ "Cumacea",
                                     Species == "Brachiopoda_spp" ~ "Brachiopoda", TRUE ~ FunctionalGroup)) %>% 
  arrange(., (Species))

# Table 2. List of species
knitr::kable(sp_list, format = "latex", booktabs = TRUE, linesep = "", longtable = TRUE, col.names = c("Species","Group","Reference","Link","Observation"), caption = "List of species for the food web of the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem (MPAN-BB).") %>%
  column_spec(1:3, width="3cm") %>%
  column_spec(4:5, width="5cm") %>% 
  kable_styling(latex_options = c("hold_position", "repeat_header"), font_size = 4) %>% 
  landscape()

```