### 2023-01-25 1st_Manuscript/SupMaterial.Rmd
Taxa resolution

---
title: "Supplementary material for 'The complex network of trophic interactions in a subAntarctic Marine Protected Area'"
author: "Tomás I. Marina, Irene R. Schloss and Luciana Riccialdelli"
output: pdf_document
bibliography: "AMPBancoBurdwood.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## Methodology

### Food web construction

The following figure and tables summarize the data we used to build the network of trophic (predator-prey) interactions for the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem.
Figure 1 resumes the type and number of references reviewed. Table 1 is the complete list of trophic interactions and references that confirm each of them. Table 2 is the species list with details on trophic species cases (aggregated taxa).

```{r ref_plot, echo=FALSE, fig.cap="Summary of references used to build the network of trophic (predator-prey) interactions for the MPAN-BB ecosystem."}

# Load packages
library(tidyverse)
library(ggplot2)

# Load data
load("../data/cleaned-data_jan23.rda")

# Discard low resolution interactions
int_good_res <- int_raw %>% 
  filter(!str_detect(Prey, "\\*$")) %>% 
  filter(!str_detect(Predator, "\\*$")) %>% 
  dplyr::select(Prey, Predator, PreyGroup, PredGroup, Reference) %>% 
  distinct(Prey, Predator, .keep_all = TRUE)

# Figure 1. Plot resuming references
biblio <- data.frame(unique(int_good_res$Reference))
colnames(biblio) <- "Reference"
biblio <- biblio %>% 
  mutate(Type = case_when(str_detect(Reference, "pers.com.") ~ "PersComm",
                          str_detect(Reference, "etal") ~ "Article",
                          str_detect(Reference, "&") ~ "Article",
                          str_detect(Reference, "Encyclopedia of Life") ~ "DataBase",
                          str_detect(Reference, "GloBI") ~ "DataBase",
                          str_detect(Reference, "GATEWAy") ~ "DataBase",
                          str_detect(Reference, "PhDTesis") ~ "Thesis",
                          TRUE ~ "Article"))
ggplot(biblio) +
  geom_bar(aes(x = Type, fill = Type)) +
  theme_bw() +
  labs(x = "Reference", y = "Count") +
  annotate("text", x = Inf, y = Inf, label = paste("Total = ", nrow(biblio), sep = ""), vjust=1, hjust=1)

```

\newpage

```{r tableS1, echo=FALSE, results='asis'}

# Load packages
library(knitr)
library(dplyr)
library(kableExtra)
library(pander)

# Load data
load("../data/cleaned-data_jan23.rda")

# Discard low resolution interactions
int_good_res <- int_raw %>% 
  filter(!str_detect(Prey, "\\*$")) %>% 
  filter(!str_detect(Predator, "\\*$")) %>% 
  dplyr::select(Prey, Predator, Reference, Reference_link) %>% 
  distinct(Prey, Predator, .keep_all = TRUE)

# Table 1. List of interactions
knitr::kable(int_good_res %>% arrange(., (Predator)), booktabs = TRUE, format = "latex", longtable = T, col.names = c("Prey","Predator","Reference", "Link"), caption = "List of predator-prey (trophic) interactions for the food web of the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem. Alphabetically ordered by predator.") %>% 
  column_spec(1:3, width="2cm") %>% 
  column_spec(4, width="14cm") %>% 
  kable_styling(latex_options = c("hold_position","repeat_header"), font_size = 4) %>% 
  landscape()

# pandoc.table(int_good_res %>% arrange(., (Predator)) %>% mutate(Reference_link = paste('\\tiny', Reference_link)) %>% rename("Link"="Reference_link"), style="multiline", split.tables = Inf, justify = c("left","left","left","right"),
#              caption = "List of predator-prey (trophic) interactions used to build the network for the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem (MPAN-BB).")

```

\newpage

```{r tableS2, echo=FALSE, results='asis'}

# Load packages
library(knitr)
library(dplyr)
library(kableExtra)

# Load data
load("../data/cleaned-data_jan23.rda")
load("../data/foodweb-data_jan23.rda")

# Tidy data
df_g <- igraph::as_data_frame(g, 'both')
sp_list <- df_g[["vertices"]] %>% 
  rename("TrophicSpecies" = "name") %>% 
  left_join(sp_raw) %>% 
  dplyr::select(TrophicSpecies, FunctionalGroup, Reference_pres, Reference_pres_link, Observation) %>% 
  rename("Species" = "TrophicSpecies", "Reference" = "Reference_pres") %>% 
  mutate(Observation = case_when(Observation %in% c("Registrado Psolus_sp; Martinez sugiere presencia de P. segregatus", "Registrado Psolus_sp; Martinez sugiere presencia de P. patagonicus", "Registrado Psolus_sp; Martinez sugiere presencia de P. antarcticus", "Lista de sp no publicada provista por Díaz de Astarloa", "Info from Sup Info (Teso_etal_2019)", "AMPNBB-II", "Abundance data available ?", "Esquema trófico", "Depredadores tope", "Copepoda", "Benthic. Not recorded", "Cirripedia") ~ "", 
                                 Species == "Campylaspis_spp" ~ "Aggregated: Campylaspis_frigida y Campylaspis_bacescui",
                                 Species == "Ophiactis_sp" ~ "Aggregated: Ophiactis_asperula y Ophiactis_spp",
                                 Species == "Brachiopoda_spp" ~ "Aggregated: Magellania_venosa, Liothyrella_uva y Terebratella_dorsata",
                                 TRUE ~ Observation),
         Reference = case_when(Species == "Bacteria" ~ "Guinder_etal_2020",
                               Species == "Necromass" ~ "Riccialdelli_etal_2017", 
                               Species == "Syllidae" ~ "Schejter_etal_2016; Bremec_etal_2019",
                               Species == "Sabellidae" ~ "Schejter_etal_2016; Bremec_etal_2019; Spinelli_etal_2020",
                               Species == "Ophiactis_sp" ~ "Schejter_etal_2016; 2020a; Brogger_pers.com.",
                               Species == "Campylaspis_spp" ~ "Doti_etal_2020",
                               Species == "Brachiopoda_spp" ~ "Schejter_etal_2016", TRUE ~ Reference),
         FunctionalGroup = case_when(Species == "Campylaspis_spp" ~ "Cumacea",
                                     Species == "Brachiopoda_spp" ~ "Brachiopoda", TRUE ~ FunctionalGroup)) %>% 
  mutate(TaxLevel = case_when(str_detect(Species, "_") | str_detect(Species, "_spp") ~ "Species",
                              str_detect(Species, "_sp") ~ "Genus",
                              str_detect(Species, "dae$") ~ "Family",
                              Species %in% c("Detritus","Necromass") ~ "NA",
                              Species %in% c("Bacteria","Foraminifera") ~ "Phylum",
                              Species %in% c("Calcarea","Copepodite","Hexactinellida") ~ "Class",
                              Species %in% c("Halichondrida","Haplosclerida","Poecilosclerida","Tetractinellida") ~ "Order",
                              TRUE ~ "Other")) %>% 
  dplyr::select(Species, TaxLevel, FunctionalGroup, Reference, Reference_pres_link, Observation) %>% 
  arrange(., (Species))

# Food web resolution
res_sp <- sum(sp_list$TaxLevel == "Species") / nrow(sp_list)

# Table 2. List of species
knitr::kable(sp_list, format = "latex", booktabs = TRUE, linesep = "", longtable = TRUE, col.names = c("Species","Taxonomy","Group","Reference","Link","Observation"), caption = "List of species for the food web of the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem. Alphabetically ordered by species.") %>%
  column_spec(1:3, width="2cm") %>%
  column_spec(4, width="3cm") %>% 
  column_spec(5:6, width="5cm") %>% 
  kable_styling(latex_options = c("hold_position", "repeat_header"), font_size = 4) %>% 
  landscape()

```

\newpage

## Results

Table 3 presents an exhaustive list of the species-level properties for the food web of the MPA Namuncurá - Banco Burdwood I ecosystem.

```{r tableS3, echo=FALSE}

load("../results/summary_results_jan23.rda")

# Table 2. List of species
knitr::kable(spp_total %>% 
               mutate(TopRole = case_when(TopRole == "hubcon" ~ "netcon",
                                          TRUE ~ TopRole)) %>% 
               dplyr::select(TrophicSpecies, FunctionalGroup, TotalDegree, NumPrey, NumPred, Between, Close, TL, meanTrophicSimil, TopRole) %>% 
               arrange(., desc(TotalDegree)) %>% 
               mutate(across(where(is.numeric), round, 4)), 
             format = "latex", booktabs = TRUE, linesep = "", longtable = TRUE, col.names = c("Species","Group","Degree","NumPrey","NumPred","Between","Close","TL","TS","TR"), align = "llcccccccc", caption = "Species-level properties for the food web of the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem. Ordered by degree (total number of interactions). Between: betweenness; Close: closeness; TL: trophic level; TS: trophic similarity; TR: topological role (netcon = network connector, modcon = module connector, modhub = module hub, modspe = module specialist).") %>%
  kable_styling(latex_options = c("hold_position", "repeat_header"), font_size = 4.5)

```

\newpage

Figure 2 shows the species' topological roles in the food web graph. Species, represented by circles, are coloured by role: network connector (dark purple), module connector (light pink), module hub (light green), and module specialist (dark green). Network connector = 5 species; module connector = 35 species; module hub = 7 species; and module specialist = 332 species.

```{r toprole_graph, echo=FALSE, fig.width = 20, fig.height=20, fig.cap="Graph of the food web showing species coloured by topological role: network connector (dark purple), module connector (light pink), module hub (light green), and module specialist (dark green)."}

# Load packages
library(igraph)
library(multiweb)
library(grid)
library(dplyr)
library(ggplot2)

# Load data
load("../results/summary_results_jan23.rda")

# Topological roles
top.role.col <- hcl.colors(4, palette = "Purple-Green")
TRColor <- setNames(top.role.col, levels(as.factor(V(g_up)$TopRole)))
TRColor <- as.data.frame(TRColor) %>% 
  mutate(TopRole = c("hubcon", "modcon", "modhub", "modspe"))
#show_col(top.role.col)  # shows color & code
data_TRColor <- spp_total %>% 
  dplyr::select(TopRole, everything()) %>% 
  left_join(TRColor)
g_up_c <- g_up %>% 
  set_vertex_attr("TRColor", value = data_TRColor$TRColor)

# Number of species by Role
TR_table <- spp_total %>% 
  group_by(TopRole) %>% 
  summarise(n = n()) %>%
  mutate(freq = n / sum(n)) %>% 
  mutate(TopRole = case_when(TopRole == "hubcon" ~ "Network connector",
                             TopRole == "modcon" ~ "Module connector",
                             TopRole == "modhub" ~ "Module hub",
                             TopRole == "modspe" ~ "Module specialist"))

# Filter network connectors ('hubcon') & module connectors ('modcon'). Violet & pink respectively
sp_con <- spp_total %>% 
  filter(TopRole == "hubcon" | TopRole == "modcon")

# Move to a new page
grid.newpage()
# Arrange figure
par(cex.axis=1.5, cex.lab=1.5)
plot_troph_level(g_up_c, vertexLabel = F, modules = F, vertex.color = V(g_up_c)$TRColor, ylab = "Trophic level")

```