### 2023-03-28 MS_ToSubmit/MS_Burdwood_foodweb.tex
Add LS & IR minor comments

\documentclass[preprint, 3p,
authoryear]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%

\usepackage[hyphens]{url}

  \journal{MEPS, Oikos, Ecological Applications, J of Applied Ecology
(?)} % Sets Journal name

\usepackage{lineno} % add
  \linenumbers % turns line numbering on

\usepackage{graphicx}
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={The complex network of trophic interactions in a subAntarctic Marine Protected Area},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

\begin{frontmatter}

  \title{The complex network of trophic interactions in a subAntarctic
Marine Protected Area}
    \author[]{Tomás I. Marina\textsuperscript{a}%
  %
  }
   \ead{tomasimarina@gmail.com} 
    \author[]{Irene R. Schloss\textsuperscript{a,b}%
  %
  }

    \author[]{Luciana Riccialdelli\textsuperscript{a}%
  %
  }

    \author[]{To be continued(?)\textsuperscript{c}%
  %
  }

      \affiliation[1]{Centro Austral de Investigaciones Cientificas
(CADIC-CONICET), Argentina}
    \affiliation[2]{Instituto Antartico Argentino (IAA), Argentina}
    \affiliation[3]{INIDEP, UBA-CONICET, IIMyC-CONICET, MACN-CONICET,
UNC-CONICET}
    \cortext[cor1]{Corresponding author}

  \begin{abstract}
  The abstract goes here.
  \end{abstract}
    \begin{keyword}
    Food web \sep Complexity \sep Structure \sep Marine Protected
Area \sep 
    Southwest Atlantic
  \end{keyword}

 \end{frontmatter}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The evidence of benefits provided by Marine Protected Areas (MPAs) as
well as the urgent need for ocean protection have driven an
unprecedented increase in the number of MPAs worldwide in recent years
(Roberts et al., 2017; Sala et al., 2018). Globally, the total area of
the world ocean designated under marine protection adds up to nearly
29,600,000 km\textsuperscript{2}, distributed across nearly 18,444 MPAs
and covering 8.16\% of the ocean's surface (IUCN \& UNEP-WCMC, 2023),
and therefore approaching the 10\% goal of the Convention of Biological
Diversity (CBD, 2004). Despite this progress, recent reports have shown
that actual protection has been overestimated because it includes areas
that are not yet effectively protected (only declared) as well as areas
that allow significant extractive activities (Sala et al., 2018).

In the sub-Antarctic region, the level of ocean protection is mainly
associated to oceanic islands, such as the South Georgia and South
Sandwich, Bouvet, Prince Edward, and Macquarie islands (IUCN \&
UNEP-WCMC, 2023). Interestingly, the case of the MPAs Namuncurá - Banco
Burdwood I and II (MPA N-BB), which is the focus of this work, is unique
since these MPAs are associated to a submarine plateau and its southern
adjacent deep slope region, respectively (Falabella, 2017). Added to
this, such MPAs are part of a network of protected areas in the
sub-Antarctic (jointly with MPA Yaganes), that aims to protect and
strengthen the functioning of this southern region to maintain global
ocean health.

Many of these MPAs focus on the presence of particularly vulnerable,
keystone, or charismatic species, the presence of large numbers (or
proportions) of endemic species, and/or the presence of high
biodiversity across taxonomic levels (Hogg et al., 2016). Indeed, the
MPA N-BB was created to protect a sensitive and biodiverse benthic
habitat that was barely known (Falabella, 2017; L. Schejter et al.,
2016). The benthic community is featured by a high biomass of vulnerable
and fragile species (mainly Poriferaand Cnidaria) that meet the
characteristics of vulnerable marine ecosystems (Schejter \& Albano,
2021), here defined as sites that present taxa with a minimum biomass of
10 kg per 1200 m\textsuperscript{2} (Commission for the Conservation of
Antarctic Marine Living Resources, 2009). Also, it provides habitat to
several small-sized species (López-Gappa et al., 2018; L. Schejter \&
Bremec, 2019; Sirito, 2019), and has an important role in the life
history of fishes as a food source, refuge and nursery area (Delpiani et
al., 2020; García Alonso et al., 2018; Troccoli et al., 2020). The
maintenance of this singular community is related to local and regional
oceanographic processes, including the circulation of the rich Malvinas
(Falkland) current in the area (Guerrero et al., 1999; A. R. Piola \&
Gordon, 1989) and the upwelling and mixing phenomena (Matano et al.,
2019). The input of nutrients coming from the Malvinas (Falkland)
current also supports a diverse plankton community (Guinder et al.,
2020). Overall, 811 benthic and plankton species have been identified
for the MPA N-BB ecosystem, where 349 were reported for the first time
for the area (APN, 2022). However, identifying the main species for
maintenance of ecosystem services and health as well as for management
and conservation is essential.

Recently, the structure of the southwestern South Atlantic Ocean has
been proposed to be under a `wasp-waist' control, meaning that the
structure and dynamics of the ecosystem is regulated primarily by
mid-trophic level species (e.g., fishes, crustaceans) (Padovani et al.,
2012; Riccialdelli et al., 2020; Saporiti et al., 2015). In particular,
the ecosystem of the MPA N-BB shows a more pronounced `wasp-waist'
structure than other sub-Antarctic areas, such as the Atlantic coast of
Tierra del Fuego. The Fuegian sprat \emph{Sprattus fuegensis} and
longtail southern cod \emph{Patagonotothen ramsayi} are considered the
most plausible `wasp-waist' species (Riccialdelli et al., 2020).

High-latitude marine ecosystems, such as the MPA N-BB, are characterized
by an unexpected complexity in terms of biodiversity and ecological
interactions (Cordone et al., 2020; Day et al., 2013; Kortsch et al.,
2019; Trathan et al., 2021). Although there is a robust knowledge about
the complexity considering the richness of the benthic and plankton
communities in the MPA N-BB ecosystem (APN, 2022; Guinder et al., 2020;
L. Schejter et al., 2016, 2020), there is a need for a better
understanding of species interactions' complexity and structure. This
aspect can be tackled by analyzing one of the most-frequent
relationships occurring between species: the predator-prey interaction
(Bascompte, 2009). The sum of predator-prey or trophic interactions of a
particular region is referred to as a food web, which represents the
roadmap for matter and energy flow in an ecosystem. In recent years,
network approaches have been successfully applied to study complex
high-latitude marine ecosystems, improving our knowledge on structure,
functioning, and response to environmental/anthropogenic changes
(Cordone et al., 2018; Funes et al., 2022; Kortsch et al., 2015; Marina
et al., 2023). Among anthropogenic threats, it is worth mentioning that
contaminants like mercury and microplastics have been recently reported
as important threats for the MPA N-BB region (Cossi et al., 2021; Di
Mauro et al., 2022; Fioramonti et al., 2022).

In the present work, we present the first, detailed analysis of the
network of predator-prey interactions, hereafter food web, for the MPA
N-BB ecosystem. For this, we applied a network approach to a highly
resolved food web. The objective was twofold: characterize the food web
in terms of complexity and structure, and identify the species' role in
such network framework.

\hypertarget{methodology}{%
\section{Methodology}\label{methodology}}

\hypertarget{study-area}{%
\subsection{Study area}\label{study-area}}

The MPAs Namuncurá - Banco Burdwood I and II, created by National Laws
26.875 and 27.490, comprise a shallow submarine plateau called Burdwood
Bank (BB) and a deep slope that reaches 4000 m in depth, N-BB I and N-BB
II, respectively (APN, 2023). They are located 150 km east of isla de
los Estados and 200 km south from Malvinas/Falkland Islands (Figure 1).
The MPA N-BB I comprises nearly 34,000 km\textsuperscript{2}
circumscribed by the 200 m isobath, between 54º--55ºS and 56º--62ºW,
with a slight slope extended nearly 370 km east--west. Physical features
in the BB are fairly stable, with salinity averaging 34 all year round
and temperature ranging between 4 and 8ºC (Acha et al., 2004; Guerrero
et al., 1999; A. Piola \& Falabella, 2009). The BB is surrounded by
steep flanks of up to 4000 m depth through which strong currents
circulate (Matano et al., 2019; A. R. Piola \& Gordon, 1989; Reta,
2014). Such deep slope form the basis of N-BB II that protects about
32,000 km\textsuperscript{2}. Intense upwelling and mixing occur over
it, entraining deep nutrient-rich waters into the photic layer (Matano
et al., 2019; A. Piola \& Falabella, 2009), and resulting in a fairly
homogeneous water column both spatially and temporally (Glorioso \&
Flather, 1995; Guerrero et al., 1999; Matano et al., 2019).

Given the evidence collected during several research cruises about the
oceanographic and ecological processes connecting MPAs N-BB I and II
(references in APN, 2022), a joint management plan was recently proposed
(APN, 2022).

\begin{figure}
\includegraphics[width=1\linewidth]{MPABurdwood_map} \caption{Marine Protected Areas Namuncurá - Banco Burdwood I (MNR and MNP, northern section) and II (MNR and SMNR, southern section). Acronyms indicate categories according to the management plan: MNR - Marine National Reserve, MNP - Marine National Park and RMNR - Restricted Marine National Reserve.}\label{fig:figure1}
\end{figure}

\hypertarget{network-construction}{%
\subsection{Network construction}\label{network-construction}}

In order to build the network of predator-prey interactions we reviewed
more than 150 references considering published articles, databases and
doctoral theses. Furthermore, we took into account personal
communications from experts belonging to the working group of the study
area (https://www.pampazul.gob.ar/tag/banco-burdwood/). The diversity of
the expertise of the authors contributing to the present study was a key
factor in enhancing the quality of the network, and inherently improved
the network representation. A list of the references used to build the
network is presented in Supplementary Material (Table S1).

Due to a lack of trophic data resolution for some species inhabiting the
study area, we followed the concept of trophic species, here defined as
aggregated groups of taxa. In most cases, we followed it when specific
data on species, in the taxonomic sense, were not available. In some
cases, we collapsed species when taxa shared the same set of predators
and prey (trophic similarity, Martinez (1991)), one of the aggregation
methods that better preserves food web functional properties (Gauzens et
al., 2013). Details about this can be found in Supplementary Material
(Table S2).

With the gathered trophic data we constructed a matrix of pairwise
interactions; a value of 1 or 0 was assigned to each element a\_ij of
the matrix depending on whether the j-species preyed or not on the
i-species. Then we transformed such matrix into an oriented graph with L
trophic interactions between S nodes or species. The orientation or
direction of the graph follows the flow of energy and matter in the
network, from prey to predator.

\hypertarget{network-analysis}{%
\subsection{Network analysis}\label{network-analysis}}

We analyzed the MPA N-BB network of trophic interactions, or food web,
at two levels: A) network, considering species and interactions of the
whole network; and B) species, considering interactions and species
related to a particular species (Table 1).

The network-level analysis aims to characterizee the food web in terms
of complexity and structure. For this we calculated several network
properties commonly used to describe empirical food webs (Pascual \&
Dunne, 2005): (1) number of species S; (2) number of interactions or
links L; (3) link density L/S; (4) connectance L/S\^{}2; (5) omnivory
Omn; and (6) small-world pattern. In order to explore the small-world
phenomenon we analyzed the characteristic path length (CPL) and the
clustering coefficient (CC). The CPL is defined as the average shortest
path length between all pairs of nodes (Watts \& Strogatz, 1998). Here,
CPL was calculated as the average number of nodes in the shortest path
\(CPL_{Min} (i,j)\) between all pairs of nodes \(S(i,j)\) in a network
averaged over \(S(S-1)/2\) nodes:

\[
CPL = \frac{2}{S(S-1)} \sum_{i = 1}^{S} \sum_{i = 1}^{S} {CPL_{Min}(i,j)}
\] The CC quantifies the local interconnectedness of the network and it
is defined as the fraction of the number of existing links between
neighbours of node \(i\) among all possible links between these
neighbours. In this study, the CC was determined as the average of the
individual clustering coefficients \(CC_i\) of all the nodes in the
network. Individual \(CC_i\) were determined as follows:

\[
CC_i = \frac{2E_i}{K_i(K_i-1)}
\] where \(E_i\) is the effective number of interactions between \(K_i\)
nearest-neighbor nodes of node \(i\) and the maximal possible number of
such interactions (Newman, 2003). With the aim of testing whether the
food web presented the small-world pattern, we compared the empirical
values of CPL and CC with those resulting from 1000 randomly generated
networks with the same size (S) and number of interactions (L),
following the method proposed by Marina et al. (2018).

Also, we estimated the (7) degree distributions for the food web, for
prey and predators, and for each functional group (e.g., Amphipoda,
Ascidiacea, Bivalvia, Fish, Marine mammals, Sea birds, etc.). The prey
and predator distributions indicate the frequency of prey among
predators, and viceversa; the functional group's degree shows the
distribution of interactions within groups.

The species-level analysis aims to describe the species' role in the
food web. For this, we considered the following properties: (1)
betweenness Btw; (2) closeness Cl; (3) trophic similarity TS; (4)
topological role TR; and (5) trophic level TL (Table 1). Topological
roles refers to the fact that food webs tend to naturally organize in
non-random, modular patterns, where modules are defined as a group of
species that interact more frequently among themselves than with species
that are not members of the module (Guimerà \& Nunes Amaral, 2005).
Species can play different roles in this respect, according to the
pattern of interactions they have within their own module and/or across
modules. We computed the topological role for each species, classified
as: module hub, species with a relatively high number of interactions,
but most within its own module; module specialist, species with
relatively few interactions and most within its own module; module
connector, species with relatively few interactions mainly between
modules; and network connector, species with high connectivity between
and within modules (Guimerà \& Nunes Amaral, 2005).

We also studied the relationship between species TL and the other
species properties by performing linear regression analyses. Thus, we
considered the TL as the dependent variable and the given property
(i.e.~betweenness, closeness, trophic similarity) as the independent
variable, and obtained the coefficients (slope and intercept) for the
linear model. Models were fitted using the least squares approach. We
also explored the topological role categories with the species TL. These
species-level properties provide a general appropriate description of
species' role in empirical complex food webs (Cirtwill et al., 2018).

All network analyses and graphs were performed in R version 4.2.2 (Team,
2022), mainly using `igraph' (Csardi \& Nepusz, 2006) and `multiweb'
(Saravia, 2022) packages. The source code and data are available at
https://github.com/TomasMarina/Banco-Burdwood.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2266}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2578}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2578}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2578}}@{}}
\caption{List of network and species-level properties analysed,
definitions, and relevant ecological implications related to food web
complexity and structure.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Name}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Definition}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Implications}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\textbf{Reference}
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Name}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Definition}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Implications}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\textbf{Reference}
\end{minipage} \\
\midrule()
\endhead
\textbf{Number of species} & Number of trophic species in a food web. &
It represents the species diversity and has implications for the
persistence of the ecosystem. & May 1973, Tilman 1996 \\
\textbf{Number of interactions} & Total number of trophic interactions
in a food web. & It represents the number of pathways along which matter
and energy can flow. & Dunne et al.~2002 \\
\textbf{Link density} & Ratio of interactions to species in a food web &
It represents the average number of interactions per species; informs
about how connected species are in the food-web. & Dunne et al.~2002 \\
\textbf{Connectance} & Proportion of potential links among species that
are actually realized. Range = 0 - 1. & It measures the probability of
interactions and is a fundamental measure of network complexity.
Connectance can be negatively or positively associated with food web
robustness, depending on the network structure (random vs non-random) or
how the strength of the interactions are distributed. & Martinez 1992 \\
\textbf{Degree distribution} & Frequency of trophic species that have k
or more interactions. & It suggests on the vulnerability of complex food
webs against random failures and intentional attacks (i.e. species
extinctions). & Albert \& Barabási 2002 \\
\textbf{Omnivory} & Species feeding on prey from more than one trophic
level. & It influences food web's stability; intermediate levels of
omnivory may stabilize it and may diffuse top-down effects thus reduce
the probability of trophic cascades. & McCann \& Hastings 1997 \\
\textbf{Small-world pattern} & A network with short path length
(distance between nodes) and high clustering coefficient (formation of
compartments) compared to random networks. & Consequences of this
structural pattern in food webs are of great importance in recognizing
evolutionary paths and the vulnerability to perturbations. & Watts \&
Strogatz 1998, Montoya \& Solé 2002 \\
\textbf{Betweenness} & Number of shortest paths going through a species.
& Species with high betweenness act as ``bridges''; if removed, would
have rapidly spreading effects in the food web. & Freeman 1978, Lai et
al.~2012 \\
\textbf{Closeness} & Number of steps required to reach every other
species from a given species. & The removal of a species with high
closeness will affect the most other species in the food web. & Freeman
1978, Lai et al.~2012 \\
\textbf{Trophic similarity} & Trophic overlap based on shared and unique
resources (prey) and consumers (predators). & It measures one of the
most important aspects of species' niches, the trophic niche, and
functional aspects of biodiversity. & Martinez 1992 \\
\textbf{Topological role} & Species role according to interactions
within and across modules (subgroups of species). & Four roles are
defined: module hub, module specialist, module connector and network
connector. Network connector and module connector roles maintain the
connectivity of the food web. & Guimera \& Nunes Amaral 2005 \\
\bottomrule()
\end{longtable}

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{network-level-properties}{%
\subsection{Network-level properties}\label{network-level-properties}}

In terms of complexity, the MPA Namuncurá - Banco Burdwood food web
(N-BB I and II) consisted of 1778 predator-prey interactions and 379
species, where 93\% of them were defined at the species taxonomical
level (Figure 2, Table S2). The food web presented a link density (e.g.,
the average number of interactions per species) of 4.69, and a
connectance of 0.01. Almost half of the consumers were omnivores (0.48),
feeding on sources at different trophic levels. The food web displayed a
small-world pattern, meaning that the path length was lower and the
clustering coefficient higher than the random networks (Table 2).

\newpage

\begin{figure}

{\centering \includegraphics{MS_Burdwood_foodweb_files/figure-latex/figure2-1}

}

\caption{Graph of the food web for the MPA Namuncurá - Banco Burdwood. Circles represent species and arrows trophic interactions. Circle diameter is relative to the number of interactions. Colour gradient indicates the trophic level.}\label{fig:figure2}
\end{figure}

\begin{table}

\caption{\label{tab:table2}Network-level properties of the MPA Namuncurá - Banco Burdwood food web. CPL: Characteristic Path Length; CC: Clustering Coefficient; SW: Small-World pattern. See table 1 for definitions and ecological relevance.}
\centering
\begin{tabular}[t]{r|r|r|r|r|r|r|l}
\hline
\textbf{Species} & \textbf{Interactions} & \textbf{Density} & \textbf{Connectance} & \textbf{Omnivory} & \textbf{CPL} & \textbf{CC} & \textbf{SW}\\
\hline
379 & 1778 & 4.69 & 0.01 & 0.48 & 2.99 & 0.08 & True\\
\hline
\end{tabular}
\end{table}

The degree distribution of the food web showed an asymmetric frequency
in the number of interactions, where most of the species had a
relatively low number of interactions and few species concentrated the
majority of them (Figure 3A). The distribution of prey among predators
showed that most of the consumers fed on a low number of prey whereas
few of them had multiple preys (Figure 3B). These were the top-five
predators in number of prey: \emph{Patagonotothen guntheri}
(Notothenioid fish, 52 prey), \emph{Patagonotothen ramsayi}
(Notothenioid fish, 50 prey), \emph{Dissostichus eleginoides}
(Notothenioid fish, 30 prey), \emph{Bathyraja brachyurops}
(Chondrichthyan, 30 prey), and \emph{Bathyraja griseocauda}
(Chondrichthyan, 28 prey). Following the same distribution pattern, few
prey presented multiple predators (Figure 3C). These were the top-five
prey in number of predators: Detritus (Non-living, 153 predators), the
three categories of Diatoms considered (benthic, centric and pennate, 72
predators on average), and species of the genus \emph{Euphausia}
(Zooplankton, 46 predators). Finally, taking into account the
interactions within each functional group, again the majority of the
interactions were concentrated in few species (Figure 3D). The most
evident species were: \emph{Themisto gaudichaudii} (Amphipoda),
\emph{Zygochlamys patagonica} (Bivalvia), \emph{Aspidostoma giganteum}
(Bryozoa), \emph{Munida gregaria} (Decapoda), \emph{Patagonotothen
ramsayi} and \emph{Patagonotothen guntheri} (bentho-pelagic fish),
\emph{Psychrolutes marmoratus} (demersal fish), and species of
\emph{Euphausia} (Zooplankton). Overall, there is an evident asymmetry
in the distribution of interactions among species at different levels in
the MPA Namuncurá - Banco Burdwood food web.

A list of the distribution of interactions per species is presented in
Supplementary Material (Table S3).

\begin{figure}

{\centering \includegraphics{MS_Burdwood_foodweb_files/figure-latex/figure3-1}

}

\caption{Degree distributions for the (A) food web, for (B) prey among predators, (C) predators among prey, and (D) for each functional group. Groups are vertically ordered by increasing trophic level; groups with less than 3 species were not plotted (e.g., pelagic fish).}\label{fig:figure3}
\end{figure}

\hypertarget{species-level-properties}{%
\subsection{Species-level properties}\label{species-level-properties}}

The majority of the species in the food web were consumers, 336 out of
379; the rest were primary producers, such as diatoms (phytoplankton),
and non-living food sources like detritus and necromass.

We found different relationships between the species trophic level (TL)
and the rest of the analysed species-level properties (Figure 4A-D). The
most evident significant relationship was with trophic similarity,
i.e.~the higher the species' TL, the lower the trophic similarity or the
higher the uniqueness in terms of trophic role (Figure 4C). Here it is
noteworthy to highlight those high-trophic level species with low values
of trophic similarity: \emph{Bathyraja macloviana} and \emph{Squalus
acanthias} (Chondrichthyans), \emph{Diplopteraster clarki} and
\emph{Pteraster sp} (echinoderms), \emph{Phalacrocorax\_atriceps} and
\emph{Eudyptes chrysocome} (sea birds), and \emph{Lagenorhynchus
cruciger} and \emph{Mesoplodon bowdoini} (marine mammals) (Table S3).

We also found a negative significant relationship between TL and
closeness, however less evident, meaning that low-TL species are
relatively closer to any other species in the food web (Figure 4B).
Species of genera \emph{Calanus} and \emph{Euphausia}, and species of
Brachiopoda, all of them with TL \textless{} 3, registered the highest
values of closeness (Table S3).

It is noteworthy that the highest values of betweenness were shown by
species of mid-TLs (3-4), meaning that those species participated in the
highest number of shortest paths between species (Figure 4A). The
following are the species with the highest values: \emph{Patagonotothen
ramsayi}, \emph{Dissostichus eleginoides}, \emph{Salilota australis}
(fishes), \emph{Doryteuthis gahi} (Cephalopoda), and
\emph{Patagonotothen guntheri} (Notothenioid fish) (Table S3).

Considering the topological role, `module specialist' species were the
most frequent and presented a wide TL range (1 - 4.77); `module hub' was
constrained to mid-TL species (2.48 - 3.92); `module connector' from low
to mid-TLs (1 - 3.86); and `network connector', the least frequent, had
all of its species in TL = 1, except for one with TL = 3.47 (Figure 4D,
see Figure S2 for species' topological roles in a food web graph
framework). Here it is important to highlight the two latter topological
roles, because they are responsible for linking modules and maintaining
the connectivity of the food web: 40 species (5 network connectors + 35
module connectors) from 20 different functional groups with a TL range =
1 - 3.86. The following were the five species with a network connector
role: Detritus, Diatoms (benthic, centric and pennate), and
\emph{Patagonotothen ramsayi}. The 35 species with a module connector
role represent these functional groups: Amphipoda, Bivalvia,
Brachiopoda, Bryozoa, Cnidaria, Cumacea, Decapoda, Detritus, Diatoms,
Dinoflagellates, Echinodermata, Fish (bentho-pelagic, demersal,
chondrichthyes), Foraminifera, Polychaeta, Porifera, Pycnogonida, Sea
birds, and Zooplankton (see Supplementary Material Table S3 for the
identity of the species).

An exhaustive list of the species-level properties is presented in
Supplementary Material (Table S3).

\begin{figure}

{\centering \includegraphics{MS_Burdwood_foodweb_files/figure-latex/figure4-1}

}

\caption{Species-level properties by trophic level: (A) betweenness, (B) closeness, (C) trophic similarity, and (D) topological role. Each point represents a species. Linear regressions for betweenness ($y = 72.48x - 111.98, R^2 = 0.04, p-value < 3.38e-05$), closeness ($y = 5.78e-06x - 9.37e-04, R^2 = -0.0005, p-value = 0.37$) and trophic similarity ($y = -0.01x - 0.11, R^2 = 0.07, p-value = 6.76e-08$). Note that for A, B and C panels only species with TLs equal or greater than 2 were considered.}\label{fig:figure4}
\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{the-food-web-of-the-mpa-namuncuruxe1---banco-burdwood-ecosystem}{%
\subsection{The food web of the MPA Namuncurá - Banco Burdwood
ecosystem}\label{the-food-web-of-the-mpa-namuncuruxe1---banco-burdwood-ecosystem}}

The food web of the MPA Namuncurá - Banco Burdwood ecosystem analysed in
this study is one of the most highly-resolved networks of trophic
interactions ever studied not only for a high-latitude marine ecosystem
but also for a marine protected area worldwide. It is of paramount
importance to consider the complexity of species interactions in order
to gain insights into the structure and functioning of the ecosystem,
since the aggregation of species might mask food web properties and/or
produce type II errors (false positive) (Gauzens et al., 2013; Martinez,
1993).

Food web connectance is a feature that resumes the complexity of the
network, but more importantly, it is an emergent property of pairwise
species interactions (Poisot \& Gravel, 2014). It contains information
regarding how interactions within an ecological network are distributed,
and predicts reasonably well key dynamical properties of ecological
networks (Dunne et al., 2002a). Complex marine food webs (i.e.~with more
than 25 trophic species) show connectance values ranging from 0.01 -
0.27 (Marina et al., 2018). In particular, food webs from high-latitude
regions tend to exhibit a connectance closer to the minimum (between
0.01 and 0.05) (Kortsch et al., 2015; Rodriguez et al., 2022; Santana et
al., 2013). Whether food webs display a low or a high connectance helps
to better comprehend ecosystem's synthetic properties like robustness.
In this sense, empirical analyses support the notion that
highly-connected ecological networks are robust against external
perturbations such as the introduction of new (e.g., invasive) species
(Smith-Ramesh et al., 2017) as well as species removal (e.g., local
extinction) (Dunne et al., 2002b; Montoya \& Solé, 2003). The
connectance of the food web of the MPA Namuncurá - Banco Burdwood (0.01)
is one of the lowest reported so far for these regions; in particular,
it appears to be five times lower than that of Beagle Channel, an
adjacent area (Rodriguez et al., 2022).

The degree distribution or the distribution of number of interactions
per species is the core for the structure of species interactions, which
influences the opportunities for multiple species to persist in the long
term and, therefore, their coexistence (Godoy et al., 2018). The food
web for the MPA N-BB presents an asymmetric degree distribution. This
pattern was identified at different levels of analysis: food web,
predator, prey, and functional group. Such asymmetry is a well-known
feature in empirical complex food webs in particular (Dunne et al.,
2002a; Montoya \& Solé, 2003; Stouffer et al., 2005), and has received
great attention in complex networks in general (Albert \& Barabási,
2002; Newman, 2003). The degree distribution affects the resilience of
complex food webs against random failures and intentional attacks: food
webs showing right-skewed distributions, like the one described in this
study, are more vulnerable to the removal of the most connected species
or hubs, with the potential of producing secondary extinctions and a
catastrophic fragmentation of the network (Albert et al., 2000; Dunne et
al., 2002b; Eklöf \& Ebenman, 2006).

It is suggested that the small-world pattern, i.e., a network with short
path length and high clustering coefficient is not frequent in complex
marine food webs, mainly due to a low clustering coefficient compared to
random networks (Dunne et al., 2002c; Marina et al., 2018). However, the
food web of the MPA N-BB does display a small-world pattern.
Consequences of this could be of great importance in recognizing
evolutionary paths and the vulnerability to perturbations (Montoya \&
Solé, 2002). On the one hand, a short path length implies a rapid spread
of an impact (e.g., contaminant, population fluctuation, local
extinction) throughout the network and, at the same time, a more
potentially adaptive dynamics in face of external perturbations (Montoya
\& Solé, 2002; Williams et al., 2002). On the other hand, a high
clustering coefficient indicates the formation of subnetworks composed
only by the neighbours of particular species. This translates into a
greater resistance of the network due to the confinement of
perturbations mainly within subnetworks and not spreading between them
(Heer et al., 2020; Kortsch et al., 2019). Overall, a small-world
topology provides ecological networks with a greater resilience and
resistance (Bornatowski et al., 2017; Dormann et al., 2017).

Omnivory acts as a buffer to changes as the ecosystem presents
alternative energy pathways in the face of perturbations; i.e., reducing
the risk of cascading extinctions following primary loss of species
(Borrvall et al., 2000). This is based on the fact that omnivores are
species able to adapt faster and to a wider range of environmental
conditions by changing their foraging habits to feed on the most
abundant prey (Fagan, 1997). Furthermore, omnivory can be analysed from
the interaction point of view: theoretical studies have identified
omnivorous interactions as a possible candidate for a keystone
interaction, sensu Kadoya et al. (2018), highlighting the importance of
omnivory inn stabilizing food web dynamics (McCann \& Hastings, 1997;
Neutel et al., 2002). The high proportion of omnivory in the food web of
the MPA N-BB suggests that the network might be robust to variations in
prey abundances, which could increase food web's persistence and
stability (Stouffer \& Bascompte, 2010).

To resume, the food web of the MPA Namuncurá - Banco Burdwood presents a
combination of network properties that make it unique in terms of
network resolution, complexity, and structural pattern. All this suggest
that the ecosystem might be fragile to external perturbations targeting
highly connected species. However, structural properties might provide
resilience and resistance with the final outcome of a rearrange
structure maintaining their functions.

\hypertarget{dominant-consumers-and-food-sources}{%
\subsection{Dominant consumers and food
sources}\label{dominant-consumers-and-food-sources}}

The degree distribution allows to identify important species, such as
potential keystone species (Dunne et al., 2002b; Solé \& Montoya, 2001),
generalist/specialist species, and dominant food sources (Kondoh et al.,
2010).

We have identified that the majority of the consumers in the food web of
the MPA N-BB either has a narrow diet or are specialists, while few
present a broad diet. The most evident generalist species are
\emph{Patagonotothen guntheri}, \emph{P. ramsayi}, \emph{Dissostichus
eleginoides}, \emph{Bathyraja brachyurops}, and \emph{B. griseocauda},
with more than 25 potential prey. Since these species present
mid-trophic positions in the food web (3-4), acting as predator and
prey, they might be important links between lower and higher trophic
levels. This is in agreement with the sole analysis, using stable
isotopes, that exists so far for the trophic structure of the MPA N-BB
(Riccialdelli et al., 2020), and resembles other high-latitude marine
systems of the Southwest Atlantic and Arctic regions (Arkhipkin \&
Laptikhovsky, 2013; Kortsch et al., 2019). The importance of these
particular generalist species also arises since they feed on the benthic
and pelagic habitats (Covatti Ale et al., 2022; Fischer et al., 2022;
Troccoli et al., 2020), linking these realms and contributing to the
vertical carbon flow.

On the other hand, few prey are consumed by many predators in the food
web of the MPA N-BB. This suggests that there are dominant food sources
from which most consumers depend on, and from where the ecosystem energy
is being transferred to the upper trophic levels. The most demanded
source we identified in this study (i.e.~detritus) supports the abundant
benthic community of filter-feeders (L. Schejter et al., 2016).
Furthermore, we found that the second and third-most consumed prey were
diatoms and species of \emph{Euphausia}, respectively, which are
essential sources for the diverse zooplankton community (Spinelli et
al., 2020), mid-TL consumers like the Fuegian sprat \emph{Sprattus
fuegensis} (Padovani et al., 2021) and \emph{Patagonotothen ramsayi}
(Fischer et al., 2022), and top predators (Raya Rey, 2022; Riccialdelli,
2022).

\hypertarget{species-role-related-to-their-trophic-level}{%
\subsection{Species' role related to their trophic
level}\label{species-role-related-to-their-trophic-level}}

Describing species' roles in food webs provides a toolbox with which to
assess the significance of species in terms of community's functioning
and overall stability (Cirtwill et al., 2018; Thébault \& Fontaine,
2010). We used a range of descriptors to characterise the dynamic and
multifaceted nature of the species forming the MPA N-BB food web.

Closeness and betweenness are defined as ``meso-scale'' properties
because they consider direct and indirect interactions, therefore
describing the focal species' ability to influence the rest of the
species of the food web (Lai et al., 2012). Closeness quantifies how
many steps away species \(i\) is from all other species in the food web,
and is proportional to how rapidly the indirect effects of the focal
species can spread to other species in the network (Scotti \& Jordán,
2010). In the food web of the MPA N-BB, low-TL consumers arise as
important in this regard: species of the zooplankton community,
\emph{Calanus} and \emph{Euphausia}, and Brachiopoda. Any perturbation
affecting these species, such as the recently confirmed contaminants
mercury (Fioramonti et al., 2022) and microplastics (Cossi et al., 2021;
Di Mauro et al., 2022) for the area, should be of concern since it might
reach many other species in the food web. Otherwise, betweenness
measures the number of shortest paths between species, providing
information on the importance of species as ``bridges'' for energy
transfer: a species with high betweenness takes part in more food chains
and therefore affects more energy flows (Scotti \& Jordán, 2010). We
have identified the longtail southern cod \emph{Patagonotothen ramsayi}
as the most important species in this sense. Moreover, in light of our
analysis, other species like the Patagonian toothfish \emph{Dissostichus
eleginoides}, the Patagonian cod \emph{Salilota australis}, the
yellowfin notothen \emph{Patagonotothen guntheri}, and the Patagonian
longfin squid \emph{Doryteuthis gahi} should be considered as relevant
in the energy transfer in the ecosystem. All these species have a
mid-trophic position in the food web, supporting the `wasp-waist'
control hypothesis for the MPA N-BB (Riccialdelli et al., 2020).

Ecosystems with a pronounced `wasp-waist' structure are suggested to
present a high trophic redundancy, since a large proportion of species
would show similar trophic habits (Cury et al., 2000). The significant
negative relationship we found between trophic similarity and trophic
level enhances the hypothesis of functional redundancy at low and mid-TL
species compared to higher TL species for the MPA N-BB food web
(Riccialdelli et al., 2020). At the same time, our results highlight the
uniqueness in terms of trophic role for high-TL predators. Here, not
only the expected pelagic animals such as marine mammals and sea birds
arise as relevant, but also benthic vertebrate (chondrichthyans
\emph{Bathyraja macloviana} and \emph{Squalus acanthias}) and
invertebrate species (echinoderms \emph{Diplopteraster clarki} and
\emph{Pteraster sp}) are noteworthy. The role that such species play in
the MPA N-BB ecosystem is unique and perturbations on them might result
in unprecedented changes at the trophic structure and functioning level.

Species' role can be also assessed in a module-based context. Among the
varying numbers of topological roles in which species can be divided,
two are remarkable: `module connector' and `network connector'. Here
again the longtail southern cod \emph{Patagonotothen ramsayi} is
emphasized as an influential species for the connectivity of the food
web, which agrees with the suggested `wasp-waist' control for the MPA
N-BB ecosystem (Riccialdelli et al., 2020). But most importantly, our
results broadens this issue by pointing out that there are other
species, belonging to a wide range of trophic positions (1 to 3.86) and
representing 20 different functional groups, that should be considered.
Thus, we propose that it is the diversity of species (benthic and
pelagic) what maintains the connectivity of the food web, therefore
contributing to the trophic structure and ecosystem's stability.

\hypertarget{caveats-and-future-perspectives}{%
\subsection{Caveats and future
perspectives}\label{caveats-and-future-perspectives}}

In construction\ldots{}

Caveats:

Seasonal variation in predators' diet due to food source availability.
Brose et al. (2019)

Spatial variation of study area: Banco Burdwood west-east gradient. In
the oceanic BB area, complex oceanographic conditions generate an
internal spatial heterogeneity, mainly along its longitudinal axis
(Matano et al., 2019), and consequently plankton communities respond
with different spatial distributions (Spinelli et al., 2020). Kortsch et
al. (2019)

Perspectives:

Incorporate habitat heterogeneity, since benthic community in the BB and
deep flanks is different (Schejter \& Albano, 2021). This might lead to
distinct food web properties, hence differences in ecosystem functioning
and response to anthropogenic and environmental perturbations. Cordone
et al. (2020)

Simulate perturbations on food web and target species: mercury transfer,
microplastic pollution, fisheries. Cossi et al. (2021); Di Mauro et al.
(2022); Fioramonti et al. (2022)

Estimation of interaction strength considering species and interaction
traits (i.e.~body size, body mass, interaction dimensionality) to gain
insights into food web stability and response to anthropogenic and
environmental perturbations. Nilsson \& McCann (2016); Cirtwill et al.
(2018); Kortsch et al. (2021)

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

We compiled information on the species and trophic diversity of the
Marine Protected Area Namuncurá - Banco Burdwood, generating an
unprecedented, well-resolved network of trophic interactions for a
sub-Antarctic ecosystem, identifying the complexity and structure of the
system, and the main species role in a network framework. Particular
properties at the network level allowed us to identify ecosystem's
vulnerability and potential response to perturbations in the presence of
highly-connected species, with a rearrange structure maintaining their
functions due to its potential resilience and resistance.

We identified several species as important in terms of different aspects
of trophic structure and functioning, and response to perturbations
(i.e.~environmental/anthropogenic changes). On one hand, we suggest that
generalist species, mostly fishes, play a crucial role in the
ecosystem's bentho-pelagic coupling process. At the same time, we
propose that other species rather than the longtail southern cod
\emph{Patagonotothen ramsayi} and the Fueguian sprat \emph{Sprattus
fuegensis} should be considered as relevant energy transfers for the
ecosystem. Finally, we argue that it is the diversity of species,
representing the benthic and pelagic habitats, what maintains the
connectivity of the food web against perturbations, therefore
contributing to the structure and stability of the ecosystem.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We are indebted to all those experts of the working group `Banco
Burdwood' who humbly provided their knowledge to enhance the quality of
the present research (add names if necessary), and special thanks to GA
Lovrich the Scientific Coordinator of the MPA Namuncurá-Burdwood Bank
and the MPA N-BB Administrators. This study was funded by Consejo
Nacional de Investigaciones Científicas y Técnicas (CONICET) and Agencia
Nacional de Promoción Científica y Tecnológica (PICT 2020-SERIEA-01617),
Argentina. This work is contribution no. XX of the MPA Namuncurá (Law
26.875).

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Acha2004}{}}%
Acha, E. M., Mianzan, H. W., Guerrero, R. A., Favero, M., \& Bava, J.
(2004). Marine fronts at the continental shelves of austral {South
America}: {Physical} and ecological processes. \emph{Journal of Marine
Systems}, \emph{44}(1), 83--105.
\url{https://doi.org/10.1016/j.jmarsys.2003.09.005}

\leavevmode\vadjust pre{\hypertarget{ref-Albert2002}{}}%
Albert, R., \& Barabási, A.-L. (2002). Statistical mechanics of complex
networks. \emph{Reviews of Modern Physics}, \emph{74}(1), 47--97.
\url{https://doi.org/10.1103/RevModPhys.74.47}

\leavevmode\vadjust pre{\hypertarget{ref-Albert2000}{}}%
Albert, R., Jeong, H., \& Barabási, A.-L. (2000). Error and attack
tolerance of complex networks. \emph{Nature}, \emph{406}(6794),
378--382. \url{https://doi.org/10.1038/35019019}

\leavevmode\vadjust pre{\hypertarget{ref-APN2022}{}}%
APN. (2022). \emph{Plan de gestión {AMP Namuncurá Banco Burdwood}}.

\leavevmode\vadjust pre{\hypertarget{ref-APN2023}{}}%
APN. (2023). \emph{Área {Marina Protegida Namuncurá Banco Burdwood}}.

\leavevmode\vadjust pre{\hypertarget{ref-Arkhipkin2013}{}}%
Arkhipkin, A., \& Laptikhovsky, V. (2013). From gelatinous to muscle
food chain: Rock cod {Patagonotothen} ramsayi recycles coelenterate and
tunicate resources on the {Patagonian Shelf}. \emph{Journal of Fish
Biology}, \emph{83}(5), 1210--1220.
\url{https://doi.org/10.1111/jfb.12217}

\leavevmode\vadjust pre{\hypertarget{ref-Bascompte2009}{}}%
Bascompte, J. (2009). Disentangling the {Web} of {Life}. \emph{Science},
\emph{325}(5939), 416--419.
\url{https://doi.org/10.1126/science.1170749}

\leavevmode\vadjust pre{\hypertarget{ref-Bergagna2019}{}}%
Bergagna, L. (2019). \emph{{Descripción de los ensambles bentónicos del
AMP Namuncurá-Banco Burdwood y estimación del contenido de carbono
-orgánico e inorgánico- allí presente.}} {[}\{Bachelor Thesis\}{]}.
Universidad Nacional de Córdoba.

\leavevmode\vadjust pre{\hypertarget{ref-Bergagna2020}{}}%
Bergagna, L., Chiesa, I., Álvarez, Y. P., Ceballos, S. G., Boy, C. C.,
Riccialdelli, L., \& Lovrich, G. A. (2020). {En el fondo el Banco es
bueno}. \emph{La Lupa. Colección Fueguina de Divulgación Científica},
\emph{16}, 20--24.

\leavevmode\vadjust pre{\hypertarget{ref-Bornatowski2017}{}}%
Bornatowski, H., Barreto, R., Navia, A. F., \& de Amorim, A. F. (2017).
Topological redundancy and {``small-world''} patterns in a food web in a
subtropical ecosystem of {Brazil}. \emph{Marine Ecology}, \emph{38}(2),
e12407. \url{https://doi.org/10.1111/maec.12407}

\leavevmode\vadjust pre{\hypertarget{ref-Borrelli2014}{}}%
Borrelli, J. J., \& Ginzburg, L. R. (2014). Why there are so few trophic
levels: {Selection} against instability explains the pattern. \emph{Food
Webs}, \emph{1}(1), 10--17.
\url{https://doi.org/10.1016/j.fooweb.2014.11.002}

\leavevmode\vadjust pre{\hypertarget{ref-Borrvall2000}{}}%
Borrvall, C., Ebenman, B., \& Tomas Jonsson, T. J. (2000). Biodiversity
lessens the risk of cascading extinction in model food webs.
\emph{Ecology Letters}, \emph{3}(2), 131--136.
\url{https://doi.org/10.1046/j.1461-0248.2000.00130.x}

\leavevmode\vadjust pre{\hypertarget{ref-Brickle2009}{}}%
Brickle, P., Arkhipkin, A. I., Laptikhovsky, V., Stocks, A., \& Taylor,
A. (2009). Resource partitioning by two large planktivorous fishes
{Micromesistius} australis and {Macruronus} magellanicus in the
{Southwest Atlantic}. \emph{Estuarine, Coastal and Shelf Science},
\emph{84}(1), 91--98. \url{https://doi.org/10.1016/j.ecss.2009.06.007}

\leavevmode\vadjust pre{\hypertarget{ref-Brose2019}{}}%
Brose, U., Archambault, P., Barnes, A. D., Bersier, L.-F., Boy, T.,
Canning-Clode, J., Conti, E., Dias, M., Digel, C., Dissanayake, A.,
Flores, A. A. V., Fussmann, K., Gauzens, B., Gray, C., Häussler, J.,
Hirt, M. R., Jacob, U., Jochum, M., Kéfi, S., \ldots{} Iles, A. C.
(2019). Predator traits determine food-web architecture across
ecosystems. \emph{Nature Ecology \& Evolution}, \emph{3}(6), 919--927.
\url{https://doi.org/10.1038/s41559-019-0899-x}

\leavevmode\vadjust pre{\hypertarget{ref-CBD2004}{}}%
CBD. (2004). \emph{Secretariat of the convention on biological
diversity: {Technical} advice on the establishment and management of a
national system of marine and coastal protected areas, {Technical
Series} number 13.}

\leavevmode\vadjust pre{\hypertarget{ref-Cirtwill2018}{}}%
Cirtwill, A. R., Dalla Riva, G. V., Gaiarsa, M. P., Bimler, M. D.,
Cagua, E. F., Coux, C., \& Dehling, D. M. (2018). A review of species
role concepts in food webs. \emph{Food Webs}, \emph{16}, e00093.
\url{https://doi.org/10.1016/j.fooweb.2018.e00093}

\leavevmode\vadjust pre{\hypertarget{ref-CommissionfortheConservationofAntarcticMarineLivingResources2009}{}}%
Commission for the Conservation of Antarctic Marine Living Resources.
(2009). \emph{Vulnerable {Marine Ecosystem} taxa identification guide}.

\leavevmode\vadjust pre{\hypertarget{ref-Cordone2018}{}}%
Cordone, G., Marina, T. I., Salinas, V., Doyle, S. R., Saravia, L. A.,
\& Momo, F. R. (2018). Effects of macroalgae loss in an {Antarctic}
marine food web: Applying extinction thresholds to food web studies.
\emph{PeerJ}, \emph{6}, e5531. \url{https://doi.org/10.7717/peerj.5531}

\leavevmode\vadjust pre{\hypertarget{ref-Cordone2020}{}}%
Cordone, G., Salinas, V., Marina, T. I., Doyle, S. R., Pasotti, F.,
Saravia, L. A., \& Momo, F. R. (2020). Green vs brown food web:
{Effects} of habitat type on multidimensional stability proxies for a
highly-resolved {Antarctic} food web. \emph{Food Webs}, \emph{25},
e00166. \url{https://doi.org/10.1016/j.fooweb.2020.e00166}

\leavevmode\vadjust pre{\hypertarget{ref-Cossi2021}{}}%
Cossi, P. F., Ojeda, M., Chiesa, I. L., Rimondino, G. N., Fraysse, C.,
Calcagno, J., \& Pérez, A. F. (2021). First evidence of microplastics in
the {Marine Protected Area Namuncurá} at {Burdwood Bank}, {Argentina}: A
study on {Henricia} obesa and {Odontaster} penicillatus
({Echinodermata}: {Asteroidea}). \emph{Polar Biology}, \emph{44}(12),
2277--2287. \url{https://doi.org/10.1007/s00300-021-02959-5}

\leavevmode\vadjust pre{\hypertarget{ref-CovattiAle2022}{}}%
Covatti Ale, M., Fischer, L., Deli Antoni, M., Diaz de Astarloa, J. M.,
\& Delpiani, G. (2022). Trophic ecology of the yellowfin notothen,
{Patagonotothen} guntheri ({Norman}, 1937) at the {Marine Protected Area
Namuncurá-Burdwood Bank}, {Argentina}. \emph{Polar Biology},
\emph{45}(4), 549--558. \url{https://doi.org/10.1007/s00300-022-03011-w}

\leavevmode\vadjust pre{\hypertarget{ref-Csardi2006}{}}%
Csardi, \& Nepusz. (2006). \emph{The igraph software package for complex
network research}.

\leavevmode\vadjust pre{\hypertarget{ref-Cury2000}{}}%
Cury, P., Bakun, A., Crawford, R. J. M., Jarre, A., Quiñones, R. A.,
Shannon, L. J., \& Verheye, H. M. (2000). Small pelagics in upwelling
systems: Patterns of interaction and structural changes in
{``wasp-waist''} ecosystems. \emph{ICES Journal of Marine Science},
\emph{57}(3), 603--618. \url{https://doi.org/10.1006/jmsc.2000.0712}

\leavevmode\vadjust pre{\hypertarget{ref-Day2013}{}}%
Day, R. H., Weingartner, T. J., Hopcroft, R. R., Aerts, L. A. M.,
Blanchard, A. L., Gall, A. E., Gallaway, B. J., Hannay, D. E., Holladay,
B. A., Mathis, J. T., Norcross, B. L., Questel, J. M., \& Wisdom, S. S.
(2013). The offshore northeastern {Chukchi Sea}, {Alaska}: {A} complex
high-latitude ecosystem. \emph{Continental Shelf Research}, \emph{67},
147--165. \url{https://doi.org/10.1016/j.csr.2013.02.002}

\leavevmode\vadjust pre{\hypertarget{ref-Delpiani2020}{}}%
Delpiani, S. M., Bruno, D. O., Vazquez, D. M., Llompart, F., Delpiani,
G. E., Fernández, D. A., Rosso, J. J., Mabragaña, E., \& Díaz de
Astarloa, J. M. (2020). Structure and distribution of fish assemblages
at {Burdwood Bank}, the first {Sub-Antarctic Marine Protected Area}
{``{Namuncurá}''} in {Argentina} ({Southwestern Atlantic Ocean}).
\emph{Polar Biology}, \emph{43}(11), 1783--1793.
\url{https://doi.org/10.1007/s00300-020-02744-w}

\leavevmode\vadjust pre{\hypertarget{ref-DiMauro2022}{}}%
Di Mauro, R., Castillo, S., Pérez, A., Iachetti, C. M., Silva, L.,
Tomba, J. P., \& Chiesa, I. L. (2022). Anthropogenic microfibers are
highly abundant at the {Burdwood Bank} seamount, a protected
sub-{Antarctic} environment in the {Southwestern Atlantic Ocean}.
\emph{Environmental Pollution}, \emph{306}, 119364.
\url{https://doi.org/10.1016/j.envpol.2022.119364}

\leavevmode\vadjust pre{\hypertarget{ref-Dormann2017}{}}%
Dormann, C. F., Fründ, J., \& Schaefer, H. M. (2017). Identifying
{Causes} of {Patterns} in {Ecological Networks}: {Opportunities} and
{Limitations}. \emph{Annual Review of Ecology, Evolution, and
Systematics}, \emph{48}, 559--584.
\url{https://doi.org/10.1146/annurev-ecolsys-110316-022928}

\leavevmode\vadjust pre{\hypertarget{ref-Dunne2002a}{}}%
Dunne, J. A., Williams, R. J., \& Martinez, N. D. (2002a). Food-web
structure and network theory: {The} role of connectance and size.
\emph{Proceedings of the National Academy of Sciences}, \emph{99}(20),
12917--12922. \url{https://doi.org/10.1073/pnas.192407699}

\leavevmode\vadjust pre{\hypertarget{ref-Dunne2002}{}}%
Dunne, J. A., Williams, R. J., \& Martinez, N. D. (2002b). Network
structure and biodiversity loss in food webs: Robustness increases with
connectance. \emph{Ecology Letters}, \emph{5}(4), 558--567.
\url{https://doi.org/10.1046/j.1461-0248.2002.00354.x}

\leavevmode\vadjust pre{\hypertarget{ref-Dunne2002b}{}}%
Dunne, J. A., Williams, R. J., \& Martinez, N. D. (2002c). \emph{Small
{Networks} but not {Small Worlds}: {Unique Aspects} of {Food Web
Structure}}. {Santa Fe Institute}.

\leavevmode\vadjust pre{\hypertarget{ref-Eklof2006}{}}%
Eklöf, A., \& Ebenman, B. (2006). Species loss and secondary extinctions
in simple and complex model communities. \emph{Journal of Animal
Ecology}, \emph{75}(1), 239--246.
\url{https://doi.org/10.1111/j.1365-2656.2006.01041.x}

\leavevmode\vadjust pre{\hypertarget{ref-Estrada2007}{}}%
Estrada, E. (2007). Food webs robustness to biodiversity loss: {The}
roles of connectance, expansibility and degree distribution.
\emph{Journal of Theoretical Biology}, \emph{244}(2), 296--307.
\url{https://doi.org/10.1016/j.jtbi.2006.08.002}

\leavevmode\vadjust pre{\hypertarget{ref-Fagan1997}{}}%
Fagan, W. F. (1997). Omnivory as a {Stabilizing Feature} of {Natural
Communities}. \emph{The American Naturalist}, \emph{150}(5), 554--567.
\url{https://doi.org/10.1086/286081}

\leavevmode\vadjust pre{\hypertarget{ref-Falabella2017}{}}%
Falabella, V. (2017). \emph{Área {Marina Protegida Namuncurá-Banco
Burdwood}. {Contribuciones} para la línea de base y el plan de manejo}.

\leavevmode\vadjust pre{\hypertarget{ref-Fernandez2021}{}}%
Fernández, M., Rodríguez-Ruiz, M., Gelcich, S., Hiriart-Bertrand, L., \&
Castilla, J. C. (2021). Advances and challenges in marine conservation
in {Chile}: {A} regional and global comparison. \emph{Aquatic
Conservation: Marine and Freshwater Ecosystems}, \emph{31}(7),
1760--1771. \url{https://doi.org/10.1002/aqc.3570}

\leavevmode\vadjust pre{\hypertarget{ref-Fioramonti2022}{}}%
Fioramonti, N. E., Ribeiro Guevara, S., Becker, Y. A., \& Riccialdelli,
L. (2022). Mercury transfer in coastal and oceanic food webs from the
{Southwest Atlantic Ocean}. \emph{Marine Pollution Bulletin},
\emph{175}, 113365.
\url{https://doi.org/10.1016/j.marpolbul.2022.113365}

\leavevmode\vadjust pre{\hypertarget{ref-Fischer2022}{}}%
Fischer, L., Covatti Ale, M., Deli Antoni, M., Díaz de Astarloa, J. M.,
\& Delpiani, G. (2022). Feeding ecology of the longtail southern cod,
{Patagonotothen} ramsayi ({Regan}, 1913) ({Notothenioidei}) in the
{Marine Protected Area Namuncurá-Burdwood Bank}, {Argentina}.
\emph{Polar Biology}, \emph{45}(9), 1483--1494.
\url{https://doi.org/10.1007/s00300-022-03082-9}

\leavevmode\vadjust pre{\hypertarget{ref-Franco2020}{}}%
Franco, B. C., Defeo, O., Piola, A. R., Barreiro, M., Yang, H., Ortega,
L., Gianelli, I., Castello, J. P., Vera, C., Buratti, C., Pájaro, M.,
Pezzi, L. P., \& Möller, O. O. (2020). Climate change impacts on the
atmospheric circulation, ocean, and fisheries in the southwest {South
Atlantic Ocean}: A review. \emph{Climatic Change}, \emph{162}(4),
2359--2377. \url{https://doi.org/10.1007/s10584-020-02783-6}

\leavevmode\vadjust pre{\hypertarget{ref-Freeman1978}{}}%
Freeman, L. C. (1978). Centrality in social networks conceptual
clarification. \emph{Social Networks}, \emph{1}(3), 215--239.
\url{https://doi.org/10.1016/0378-8733(78)90021-7}

\leavevmode\vadjust pre{\hypertarget{ref-Funes2022}{}}%
Funes, M., Saravia, L. A., Cordone, G., Iribarne, O. O., \& Galván, D.
E. (2022). Network analysis suggests changes in food web stability
produced by bottom trawl fishery in {Patagonia}. \emph{Scientific
Reports}, \emph{12}(1), 10876.
\url{https://doi.org/10.1038/s41598-022-14363-y}

\leavevmode\vadjust pre{\hypertarget{ref-GarciaAlonso2018}{}}%
García Alonso, V. A., Brown, D., Martín, J., Pájaro, M., \& Capitanio,
F. L. (2018). Seasonal patterns of {Patagonian} sprat {Sprattus}
fuegensis early life stages in an open sea {Sub-Antarctic Marine
Protected Area}. \emph{Polar Biology}, \emph{41}(11), 2167--2179.
\url{https://doi.org/10.1007/s00300-018-2352-z}

\leavevmode\vadjust pre{\hypertarget{ref-Gauzens2013}{}}%
Gauzens, B., Legendre, S., Lazzaro, X., \& Lacroix, G. (2013). Food-web
aggregation, methodological and functional issues. \emph{Oikos},
\emph{122}(11), 1606--1615.
\url{https://doi.org/10.1111/j.1600-0706.2013.00266.x}

\leavevmode\vadjust pre{\hypertarget{ref-Glorioso1995}{}}%
Glorioso, P. D., \& Flather, R. A. (1995). A barotropic model of the
currents off {SE South America}. \emph{Journal of Geophysical Research:
Oceans}, \emph{100}(C7), 13427--13440.
\url{https://doi.org/10.1029/95JC00942}

\leavevmode\vadjust pre{\hypertarget{ref-Godoy2018}{}}%
Godoy, O., Bartomeus, I., Rohr, R. P., \& Saavedra, S. (2018). Towards
the {Integration} of {Niche} and {Network Theories}. \emph{Trends in
Ecology \& Evolution}, \emph{33}(4), 287--300.
\url{https://doi.org/10.1016/j.tree.2018.01.007}

\leavevmode\vadjust pre{\hypertarget{ref-Grilli2016}{}}%
Grilli, J., Rogers, T., \& Allesina, S. (2016). Modularity and stability
in ecological communities. \emph{Nature Communications}, \emph{7}(1),
12031. \url{https://doi.org/10.1038/ncomms12031}

\leavevmode\vadjust pre{\hypertarget{ref-Guerrero1999}{}}%
Guerrero, R. A., Baldoni, A. G., \& Benavides, H. R. (1999).
\emph{Oceanographic conditions at the southern end of the argentine
continental slope}.
https://doi.org/\url{http://10.0.64.26/handle/inidep/247}

\leavevmode\vadjust pre{\hypertarget{ref-Guimera2005}{}}%
Guimerà, R., \& Nunes Amaral, L. A. (2005). Functional cartography of
complex metabolic networks. \emph{Nature}, \emph{433}(7028), 895--900.
\url{https://doi.org/10.1038/nature03288}

\leavevmode\vadjust pre{\hypertarget{ref-Guinder2020}{}}%
Guinder, V. A., Malits, A., Ferronato, C., Krock, B., Garzón-Cardona,
J., \& Martínez, A. (2020). Microbial plankton configuration in the
epipelagic realm from the {Beagle Channel} to the {Burdwood Bank}, a
{Marine Protected Area} in {Sub-Antarctic} waters. \emph{PLOS ONE},
\emph{15}(5), e0233156.
\url{https://doi.org/10.1371/journal.pone.0233156}

\leavevmode\vadjust pre{\hypertarget{ref-Heer2020}{}}%
Heer, H., Streib, L., Schäfer, R. B., \& Ruzika, S. (2020). Maximising
the clustering coefficient of networks and the effects on habitat
network robustness. \emph{PLOS ONE}, \emph{15}(10), e0240940.
\url{https://doi.org/10.1371/journal.pone.0240940}

\leavevmode\vadjust pre{\hypertarget{ref-Hogg2016}{}}%
Hogg, O. T., Huvenne, V. A. I., Griffiths, H. J., Dorschel, B., \&
Linse, K. (2016). Landscape mapping at sub-{Antarctic South Georgia}
provides a protocol for underpinning large-scale marine protected areas.
\emph{Scientific Reports}, \emph{6}(1), 33163.
\url{https://doi.org/10.1038/srep33163}

\leavevmode\vadjust pre{\hypertarget{ref-IUCN2023}{}}%
IUCN, \& UNEP-WCMC. (2023). \emph{The world database on protected areas
({WDPA})}.

\leavevmode\vadjust pre{\hypertarget{ref-Kadoya2018}{}}%
Kadoya, T., Gellner, G., \& McCann, K. S. (2018). Potential oscillators
and keystone modules in food webs. \emph{Ecology Letters}, \emph{21}(9),
1330--1340. \url{https://doi.org/10.1111/ele.13099}

\leavevmode\vadjust pre{\hypertarget{ref-Kondoh2010}{}}%
Kondoh, M., Kato, S., \& Sakato, Y. (2010). Food webs are built up with
nested subwebs. \emph{Ecology}, \emph{91}(11), 3123--3130.
\url{https://doi.org/10.1890/09-2219.1}

\leavevmode\vadjust pre{\hypertarget{ref-Kortsch2021}{}}%
Kortsch, S., Frelat, R., Pecuchet, L., Olivier, P., Putnis, I.,
Bonsdorff, E., Ojaveer, H., Jurgensone, I., Strāķe, S., Rubene, G.,
Krūze, Ē., \& Nordström, M. C. (2021). Disentangling temporal food web
dynamics facilitates understanding of ecosystem functioning.
\emph{Journal of Animal Ecology}, \emph{90}(5), 1205--1216.
\url{https://doi.org/10.1111/1365-2656.13447}

\leavevmode\vadjust pre{\hypertarget{ref-Kortsch2019}{}}%
Kortsch, S., Primicerio, R., Aschan, M., Lind, S., Dolgov, A. V., \&
Planque, B. (2019). Food-web structure varies along environmental
gradients in a high-latitude marine ecosystem. \emph{Ecography},
\emph{42}(2), 295--308. \url{https://doi.org/10.1111/ecog.03443}

\leavevmode\vadjust pre{\hypertarget{ref-Kortsch2015}{}}%
Kortsch, S., Primicerio, R., Fossheim, M., Dolgov, A. V., \& Aschan, M.
(2015). Climate change alters the structure of arctic marine food webs
due to poleward shifts of boreal generalists. \emph{Proceedings of the
Royal Society B: Biological Sciences}, \emph{282}(1814), 20151546.
\url{https://doi.org/10.1098/rspb.2015.1546}

\leavevmode\vadjust pre{\hypertarget{ref-Lai2012}{}}%
Lai, S.-M., Liu, W.-C., \& Jordán, F. (2012). On the centrality and
uniqueness of species from the network perspective. \emph{Biology
Letters}, \emph{8}(4), 570--573.
\url{https://doi.org/10.1098/rsbl.2011.1167}

\leavevmode\vadjust pre{\hypertarget{ref-Lopez-Gappa2018}{}}%
López-Gappa, J., Liuzzi, M. G., \& Zelaya, D. G. (2018). A new genus and
species of cheilostome bryozoan associated with hermit crabs in the
subantarctic {Southwest Atlantic}. \emph{Polar Biology}, \emph{41}(4),
733--741. \url{https://doi.org/10.1007/s00300-017-2234-9}

\leavevmode\vadjust pre{\hypertarget{ref-Marina2018a}{}}%
Marina, T. I., Saravia, L. A., Cordone, G., Salinas, V., Doyle, S. R.,
\& Momo, F. R. (2018). Architecture of marine food webs: {To} be or not
be a {``small-world.''} \emph{PLOS ONE}, \emph{13}(5), e0198217.
\url{https://doi.org/10.1371/journal.pone.0198217}

\leavevmode\vadjust pre{\hypertarget{ref-Marina2023}{}}%
Marina, T. I., Saravia, L. A., \& Kortsch, S. (2023). \emph{New insights
into the {Weddell Sea} ecosystem applying a quantitative network
approach}. {EGUsphere preprint}.
\url{https://doi.org/10.5194/egusphere-2022-1518}

\leavevmode\vadjust pre{\hypertarget{ref-Martinez1992}{}}%
Martinez, N. D. (1992). Constant {Connectance} in {Community Food Webs}.
\emph{The American Naturalist}, \emph{139}(6), 1208--1218.
\url{https://doi.org/10.1086/285382}

\leavevmode\vadjust pre{\hypertarget{ref-Martinez1991}{}}%
Martinez, N. D. (1991). Artifacts or {Attributes}? {Effects} of
{Resolution} on the {Little Rock Lake Food Web}. \emph{Ecological
Monographs}, \emph{61}(4), 367--392.
\url{https://doi.org/10.2307/2937047}

\leavevmode\vadjust pre{\hypertarget{ref-Martinez1993}{}}%
Martinez, N. D. (1993). Effects of {Resolution} on {Food Web Structure}.
\emph{Oikos}, \emph{66}(3), 403--412.
\url{https://doi.org/10.2307/3544934}

\leavevmode\vadjust pre{\hypertarget{ref-Matano2019}{}}%
Matano, R. P., Palma, E. D., \& Combes, V. (2019). The {Burdwood Bank
Circulation}. \emph{Journal of Geophysical Research: Oceans},
\emph{124}(10), 6904--6926. \url{https://doi.org/10.1029/2019JC015001}

\leavevmode\vadjust pre{\hypertarget{ref-Matusevich2022}{}}%
Matusevich. (2022). \emph{Chondrichthyan fauna from the {Marine
Protected Area Namuncurá} at {Burdwood Bank}: Exploring egg nursery
grounds}. https://www.researchsquare.com.
\url{https://doi.org/10.21203/rs.3.rs-2247873/v1}

\leavevmode\vadjust pre{\hypertarget{ref-May1973}{}}%
May, R. (1973). \emph{Stability and complexity in model ecosystems}.
{Princeton University Press}.

\leavevmode\vadjust pre{\hypertarget{ref-McCann1997}{}}%
McCann, K., \& Hastings, A. (1997). Re\textendash evaluating the
omnivory\textendash stability relationship in food webs.
\emph{Proceedings of the Royal Society of London. Series B: Biological
Sciences}, \emph{264}(1385), 1249--1254.
\url{https://doi.org/10.1098/rspb.1997.0172}

\leavevmode\vadjust pre{\hypertarget{ref-Montoya2002}{}}%
Montoya, J. M., \& Solé, R. V. (2002). Small {World Patterns} in {Food
Webs}. \emph{Journal of Theoretical Biology}, \emph{214}(3), 405--412.
\url{https://doi.org/10.1006/jtbi.2001.2460}

\leavevmode\vadjust pre{\hypertarget{ref-Montoya2003}{}}%
Montoya, J. M., \& Solé, R. V. (2003). Topological properties of food
webs: From real data to community assembly models. \emph{Oikos},
\emph{102}(3), 614--622.
\url{https://doi.org/10.1034/j.1600-0706.2003.12031.x}

\leavevmode\vadjust pre{\hypertarget{ref-Neutel2002}{}}%
Neutel, A.-M., Heesterbeek, J. A. P., \& de Ruiter, P. C. (2002).
Stability in {Real Food Webs}: {Weak Links} in {Long Loops}.
\emph{Science}, \emph{296}(5570), 1120--1123.
\url{https://doi.org/10.1126/science.1068326}

\leavevmode\vadjust pre{\hypertarget{ref-Newman2003}{}}%
Newman, M. E. J. (2003). The {Structure} and {Function} of {Complex
Networks}. \emph{SIAM Review}, \emph{45}(2), 167--256.
\url{https://doi.org/10.1137/S003614450342480}

\leavevmode\vadjust pre{\hypertarget{ref-Nilsson2016}{}}%
Nilsson, K. A., \& McCann, K. S. (2016). Interaction strength
revisited\textemdash clarifying the role of energy flux for food web
stability. \emph{Theoretical Ecology}, \emph{9}(1), 59--71.
\url{https://doi.org/10.1007/s12080-015-0282-8}

\leavevmode\vadjust pre{\hypertarget{ref-Padovani2021}{}}%
Padovani, L. N., Álvarez, N., \& Farías, A. (2021). \emph{Alimentación
de la sardina fueguina ({Sprattus} fuegensis) en la región patagónica
austral durante la época reproductiva}. {INIDEP}.

\leavevmode\vadjust pre{\hypertarget{ref-Padovani2012}{}}%
Padovani, L. N., Viñas, M. D., Sánchez, F., \& Mianzan, H. (2012).
Amphipod-supported food web: {Themisto} gaudichaudii, a key food
resource for fishes in the southern {Patagonian Shelf}. \emph{Journal of
Sea Research}, \emph{67}(1), 85--90.
\url{https://doi.org/10.1016/j.seares.2011.10.007}

\leavevmode\vadjust pre{\hypertarget{ref-Pascual2005}{}}%
Pascual, M., \& Dunne, J. A. (2005). \emph{Ecological {Networks}:
{Linking Structure} to {Dynamics} in {Food Webs}}. {Oxford University
Press}.

\leavevmode\vadjust pre{\hypertarget{ref-Pastorino2019}{}}%
Pastorino, G. (2019). A new deep water gastropod of the genus
{Parabuccinum} ({Neogastropoda}: {Buccinulidae}) from southwestern
{Atlantic} waters with new data on the distribution of all species.
\emph{Marine Biodiversity}, \emph{49}(2), 913--922.
\url{https://doi.org/10.1007/s12526-018-0876-7}

\leavevmode\vadjust pre{\hypertarget{ref-Pawar2012}{}}%
Pawar, S., Dell, A. I., \& Van M. Savage. (2012). Dimensionality of
consumer search space drives trophic interaction strengths.
\emph{Nature}, \emph{486}, 485.
\url{https://doi.org/10.1038/nature11131}

\leavevmode\vadjust pre{\hypertarget{ref-Piola1989}{}}%
Piola, A. R., \& Gordon, A. L. (1989). Intermediate waters in the
southwest {South Atlantic}. \emph{Deep Sea Research Part A.
Oceanographic Research Papers}, \emph{36}(1), 1--16.
\url{https://doi.org/10.1016/0198-0149(89)90015-0}

\leavevmode\vadjust pre{\hypertarget{ref-Piola2009}{}}%
Piola, A., \& Falabella, V. (2009). \emph{El mar {Patagónico}}.
{Wildlife Conservation Society y Birdlife Internacional}.

\leavevmode\vadjust pre{\hypertarget{ref-Poisot2014}{}}%
Poisot, T., \& Gravel, D. (2014). When is an ecological network complex?
{Connectance} drives degree distribution and emerging network
properties. \emph{PeerJ}, \emph{2}, e251.
\url{https://doi.org/10.7717/peerj.251}

\leavevmode\vadjust pre{\hypertarget{ref-RayaRey2022}{}}%
Raya Rey, A. (2022). \emph{Food items of sea birds of the {Marine
Protected Area Namuncurá} - {Banco Burdwood}}.

\leavevmode\vadjust pre{\hypertarget{ref-Reta2014}{}}%
Reta, R. (2014). \emph{Oceanografía del {Banco Burdwood}: {Estado
Actual} del {Conocimiento} y {Perspectivas}}. {INIDEP}.

\leavevmode\vadjust pre{\hypertarget{ref-Riccialdelli2022}{}}%
Riccialdelli, L. (2022). \emph{Feeding items of marine mammals of the
{Marine Protected Area Namuncurá} - {Banco Burdwood}}.

\leavevmode\vadjust pre{\hypertarget{ref-Riccialdelli2020}{}}%
Riccialdelli, L., Becker, Y. A., Fioramonti, N. E., Torres, M., Bruno,
D. O., Rey, A. R., \& Fernández, D. A. (2020). Trophic structure of
southern marine ecosystems: A comparative isotopic analysis from the
{Beagle Channel} to the oceanic {Burdwood Bank} area under a wasp-waist
assumption. \emph{Marine Ecology Progress Series}, \emph{655}, 1--27.
\url{https://doi.org/10.3354/meps13524}

\leavevmode\vadjust pre{\hypertarget{ref-Roberts2017}{}}%
Roberts, C. M., O'Leary, B. C., McCauley, D. J., Cury, P. M., Duarte, C.
M., Lubchenco, J., Pauly, D., Sáenz-Arroyo, A., Sumaila, U. R., Wilson,
R. W., Worm, B., \& Castilla, J. C. (2017). Marine reserves can mitigate
and promote adaptation to climate change. \emph{Proceedings of the
National Academy of Sciences}, \emph{114}(24), 6167--6175.
\url{https://doi.org/10.1073/pnas.1701262114}

\leavevmode\vadjust pre{\hypertarget{ref-Rodriguez2022}{}}%
Rodriguez, I. D., Marina, T. I., Schloss, I. R., \& Saravia, L. A.
(2022). Marine food webs are more complex but less stable in
sub-{Antarctic} ({Beagle Channel}, {Argentina}) than in {Antarctic}
({Potter Cove}, {Antarctic Peninsula}) regions. \emph{Marine
Environmental Research}, \emph{174}, 105561.
\url{https://doi.org/10.1016/j.marenvres.2022.105561}

\leavevmode\vadjust pre{\hypertarget{ref-Sala2018}{}}%
Sala, E., Lubchenco, J., Grorud-Colvert, K., Novelli, C., Roberts, C.,
\& Sumaila, U. R. (2018). Assessing real progress towards effective
ocean protection. \emph{Marine Policy}, \emph{91}, 11--13.
\url{https://doi.org/10.1016/j.marpol.2018.02.004}

\leavevmode\vadjust pre{\hypertarget{ref-Santana2013}{}}%
Santana, C. N. de, Rozenfeld, A. F., Marquet, P. A., \& Duarte, C. M.
(2013). Topological properties of polar food webs. \emph{Marine Ecology
Progress Series}, \emph{474}, 15--26.
\url{https://doi.org/10.3354/meps10073}

\leavevmode\vadjust pre{\hypertarget{ref-Saporiti2015}{}}%
Saporiti, F., Bearhop, S., Vales, D. G., Silva, L., Zenteno, L.,
Tavares, M., Crespo, E. A., \& Cardona, L. (2015). Latitudinal changes
in the structure of marine food webs in the {Southwestern Atlantic
Ocean}. \emph{Marine Ecology Progress Series}, \emph{538}, 23--34.
\url{https://doi.org/10.3354/meps11464}

\leavevmode\vadjust pre{\hypertarget{ref-Saravia2022a}{}}%
Saravia, L. A. (2022). \emph{Multiweb: {Ecological} network analyses
including multiplex networks}.

\leavevmode\vadjust pre{\hypertarget{ref-Schejter2019}{}}%
Schejter, L., \& Bremec, C. S. (2019). Stony corals ({Anthozoa}:
{Scleractinia}) of {Burdwood Bank} and neighbouring areas, {SW Atlantic
Ocean}. \emph{Scientia Marina}, \emph{83}(3), 247--260.
\url{https://doi.org/10.3989/scimar.04863.10A}

\leavevmode\vadjust pre{\hypertarget{ref-Schejter2020}{}}%
Schejter, L., Genzano, G., Gaitán, E., Perez, C. D., \& Bremec, C. S.
(2020). Benthic communities in the {Southwest Atlantic Ocean}:
{Conservation} value of animal forests at the {Burdwood Bank} slope.
\emph{Aquatic Conservation: Marine and Freshwater Ecosystems},
\emph{30}(3), 426--439. \url{https://doi.org/10.1002/aqc.3265}

\leavevmode\vadjust pre{\hypertarget{ref-Schejter2016}{}}%
Schejter, L., Rimondino, C., Chiesa, I., Díaz de Astarloa, J. M., Doti,
B., Elías, R., Escolar, M., Genzano, G., López-Gappa, J., Tatián, M.,
Zelaya, D. G., Cristobo, J., Perez, C. D., Cordeiro, R. T., \& Bremec,
C. S. (2016). Namuncurá {Marine Protected Area}: An oceanic hot spot of
benthic biodiversity at {Burdwood Bank}, {Argentina}. \emph{Polar
Biology}, \emph{39}(12), 2373--2386.
\url{https://doi.org/10.1007/s00300-016-1913-2}

\leavevmode\vadjust pre{\hypertarget{ref-Schejter2021}{}}%
Schejter, \& Albano. (2021). Benthic communities at the marine protected
area {Namuncurá}/{Burdwood} bank, {SW Atlantic Ocean}: Detection of
vulnerable marine ecosystems and contributions to the assessment of the
rezoning process. \emph{Polar Biology}, \emph{44}(10), 2023--2037.
\url{https://doi.org/10.1007/s00300-021-02936-y}

\leavevmode\vadjust pre{\hypertarget{ref-Scotti2010}{}}%
Scotti, M., \& Jordán, F. (2010). Relationships between centrality
indices and trophic levels in food webs. \emph{Community Ecology},
\emph{11}(1), 59--67. \url{https://doi.org/10.1556/ComEc.11.2010.1.9}

\leavevmode\vadjust pre{\hypertarget{ref-MartinSirito2019}{}}%
Sirito, M. (2019). \emph{Fauna asociada a corales ({Octocorallia}) e
hidroides ({Hydrozoa}) del {Área Marina Protegida} {``{Namuncurá}''}
({Banco Burdwood}) y zonas profundas adyacentes} {[}PhD thesis{]}.
Universidad Nacional de Mar del Plata.

\leavevmode\vadjust pre{\hypertarget{ref-Smith-Ramesh2017}{}}%
Smith-Ramesh, L. M., Moore, A. C., \& Schmitz, O. J. (2017). Global
synthesis suggests that food web connectance correlates to invasion
resistance. \emph{Global Change Biology}, \emph{23}(2), 465--473.
\url{https://doi.org/10.1111/gcb.13460}

\leavevmode\vadjust pre{\hypertarget{ref-Sole2001}{}}%
Solé, R. V., \& Montoya, M. (2001). Complexity and fragility in
ecological networks. \emph{Proceedings of the Royal Society of London.
Series B: Biological Sciences}, \emph{268}(1480), 2039--2045.
\url{https://doi.org/10.1098/rspb.2001.1767}

\leavevmode\vadjust pre{\hypertarget{ref-Spinelli2020}{}}%
Spinelli, M. L., Malits, A., García Alonso, V. A., Martín, J., \&
Capitanio, F. L. (2020). Spatial gradients of spring zooplankton
assemblages at the open ocean sub-{Antarctic Namuncurá Marine Protected
Area}/{Burdwood Bank}, {SW Atlantic Ocean}. \emph{Journal of Marine
Systems}, \emph{210}, 103398.
\url{https://doi.org/10.1016/j.jmarsys.2020.103398}

\leavevmode\vadjust pre{\hypertarget{ref-Stouffer2010}{}}%
Stouffer, D. B., \& Bascompte, J. (2010). Understanding food-web
persistence from local to global scales. \emph{Ecology Letters},
\emph{13}(2), 154--161.
\url{https://doi.org/10.1111/j.1461-0248.2009.01407.x}

\leavevmode\vadjust pre{\hypertarget{ref-Stouffer2005}{}}%
Stouffer, D. B., Camacho, J., Guimerà, R., Ng, C. A., \& Nunes Amaral,
L. A. (2005). Quantitative {Patterns} in the {Structure} of {Model} and
{Empirical Food Webs}. \emph{Ecology}, \emph{86}(5), 1301--1311.
\url{https://doi.org/10.1890/04-0957}

\leavevmode\vadjust pre{\hypertarget{ref-Strogatz2001}{}}%
Strogatz, S. H. (2001). Exploring complex networks. \emph{Nature},
\emph{410}(6825), 268--276. \url{https://doi.org/10.1038/35065725}

\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam2022}{}}%
Team, R. C. (2022). \emph{R: {A Language} and {Environment} for
{Statistical Computing}}. R Foundation for Statistical Computing.

\leavevmode\vadjust pre{\hypertarget{ref-Thebault2010}{}}%
Thébault, E., \& Fontaine, C. (2010). Stability of {Ecological
Communities} and the {Architecture} of {Mutualistic} and {Trophic
Networks}. \emph{Science}, \emph{329}(5993), 853--856.
\url{https://doi.org/10.1126/science.1188321}

\leavevmode\vadjust pre{\hypertarget{ref-Tilman1996}{}}%
Tilman, D. (1996). Biodiversity: {Population Versus Ecosystem
Stability}. \emph{Ecology}, \emph{77}(2), 350--363.
\url{https://doi.org/10.2307/2265614}

\leavevmode\vadjust pre{\hypertarget{ref-Trathan2021}{}}%
Trathan, P. N., Fielding, S., Hollyman, P. R., Murphy, E. J.,
Warwick-Evans, V., \& Collins, M. A. (2021). Enhancing the ecosystem
approach for the fishery for {Antarctic} krill within the complex,
variable, and changing ecosystem at {South Georgia}. \emph{ICES Journal
of Marine Science}, \emph{78}(6), 2065--2081.
\url{https://doi.org/10.1093/icesjms/fsab092}

\leavevmode\vadjust pre{\hypertarget{ref-Troccoli2020}{}}%
Troccoli, G. H., Aguilar, E., Martínez, P. A., \& Belleggia, M. (2020).
The diet of the {Patagonian} toothfish {Dissostichus} eleginoides, a
deep-sea top predator off {Southwest Atlantic Ocean}. \emph{Polar
Biology}, \emph{43}(10), 1595--1604.
\url{https://doi.org/10.1007/s00300-020-02730-2}

\leavevmode\vadjust pre{\hypertarget{ref-Watts1998}{}}%
Watts, D. J., \& Strogatz, S. H. (1998). Collective dynamics of
{``small-world''} networks. \emph{Nature}, \emph{393}(6684), 440--442.
\url{https://doi.org/10.1038/30918}

\leavevmode\vadjust pre{\hypertarget{ref-Williams2002}{}}%
Williams, R. J., Berlow, E. L., Dunne, J. A., Barabási, A.-L., \&
Martinez, N. D. (2002). Two degrees of separation in complex food webs.
\emph{Proceedings of the National Academy of Sciences}, \emph{99}(20),
12913--12916. \url{https://doi.org/10.1073/pnas.192448799}

\leavevmode\vadjust pre{\hypertarget{ref-Woodward2005}{}}%
Woodward, G., Ebenman, B., Emmerson, M., Montoya, J. M., Olesen, J. M.,
Valido, A., \& Warren, P. H. (2005). Body size in ecological networks.
\emph{Trends in Ecology \& Evolution}, \emph{20}(7), 402--409.
\url{https://doi.org/10.1016/j.tree.2005.04.005}

\end{CSLReferences}

\end{document}