### 2024-01-11 MS_ToSubmit/RevisedSupMaterial.Rmd
Revised Sup Mat

---
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: yes
  word_document: default
# classoption: landscape
bibliography: AMPBancoBurdwood.bib
csl: "elsevier-harvard2.csl"
documentclass: article
geometry:
- vmargin=1in
- hmargin=1in
header-includes:
- \usepackage{amsmath}
- \usepackage{pdflscape,booktabs}
- \usepackage{lscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage[running]{lineno}
- \linenumbers
---

# Supplementary Material for 'The complex network of trophic interactions in a subAntarctic oceanic Marine Protected Area'

## Methodology

### Food web construction

The following figure and tables summarise the data we used to build the network of trophic (predator-prey) interactions for the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem.
Figure 2 resumes the type and number of references reviewed. Table 1 is the complete list of trophic interactions and references that confirm each of them. Table 2 is the species list with details on trophic species cases (aggregated taxa).

With the gathered trophic data, we constructed a matrix of pairwise interactions; a value of 1 or 0 was assigned to each element a_ij of the matrix depending on whether the predator j preyed or not on the prey i. Then we transformed such a matrix into an oriented graph with L trophic interactions between S nodes or species (Figure 1).

```{r, out.width = "12cm", echo = FALSE, message=FALSE, fig.align = 'center', fig.cap = "Matrix of pairwise interactions and the oriented graph (food web) derived from it. Number of species S = 9, number of interactions (L) = 15."}
library(knitr)
knitr::include_graphics("FigS1.png")
```

### Network analysis

Trophic level

Trophic levels (TL) of species were calculated using the shortweighted TL formula of @Williams2004. Shortweighted trophic level is defined as the average of the shortest TL and prey-averaged TL. Shortest TL of a consumer in a food web is equal to 1 + the shortest chain length from this consumer to any basal species [@Williams2004]. Prey averaged TL is equal to 1 + the mean TL of all consumer's trophic resources, calculated as:

\begin{equation}
TL_j = 1 + \sum_{i = 1}^{S} l_{ij} \frac{TL_i}{n_j}
\end{equation}

where $TL_j$ is the trophic level of species $j$; $S$ is the number of species in the food web; $l_{ij}$ is the connection matrix with $S$ rows and $S$ columns, in which for column $j$ and row $i$, $l_{ij}$ = 1 if species $j$ consumes species $i$ and $l_{ij}$ = 0 if not; and $n_j$ is the number of prey species in the diet of species $j$. For basal species TL = 1 given that they have no prey.

Topological roles

Beforehand, we need to calculate modularity. Modularity was calculated as the difference between realized and expected within-modules interactions, divided by the total number of interactions. We used a stochastic algorithm called “simulated annealing” [@Guimera2005], that assumes that the nodes of the same module have more links than one would expect in a random network. The modules are obtained from dividing all the nodes in the network to maximize modularity. In this way, the modularity index is defined as:

\begin{equation}
M = \sum_{s} (\frac{I_s}{L} - (\frac{d_s}{2L})^2)
\end{equation}

where $s$ is the number of modules or compartments, $I_s$ is the number of links between species in the modules, $d_s$ is the sum of degrees for all species in module $s$ and $L$ is the total number of links.

Species topological roles were calculated using the method of functional cartography [@Guimera2005]. Roles are determined according to two parameters:

a) the standardized within-module degree (dz), a z-score that reflects how well a species is connected to other species inside the module, relative to other species within its own module:

\begin{equation}
dz_i = \frac{k_{is}-k_s} {ks}
\end{equation}

where $k_{is}$ is the number of links of species $i$ within its own module $s$, and $ks$ are the average and standard deviation of $k_{is}$ over all species in $s$, respectively.

b) the participation coefficient $PC$, estimates the links distribution of species $i$ among modules:

\begin{equation}
PC_i = 1 - \sum_{S} \frac{k_{is}}{k_i}
\end{equation}

where $k_i$ is the total number of links of species $i$ and $k_{is}$ is the number of links of species $i$ to species in module $s$.
To determine the role of each species, the $dz$ − $PC$ parameters space was divided into four regions [@Kortsch2015], with two threshold values: $dz$ = 2.5 and $PC$ = 0.625. Thus, the species were classified as follows:

* Module hub ($dz$ ≥ 2.5, $PC$ < 0.625): The species has a relatively high number of links, but at least 60% within its own module.

* Module specialist ($dz$ < 2.5, $PC$ < 0.625): The species has relatively few links and most within its own module.

* Module connector ($dz$ < 2.5, $PC$ ≥ 0.625): The species has relatively few links and most between modules.

* Network connector ($dz$ ≥ 2.5, $PC$ ≥ 0.625): The species has a high connectivity between and within the modules.

\newpage
\blandscape

```{r tableS1, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Load packages
require(tidyverse)
require(knitr)
require(dplyr)
require(kableExtra)
require(pander)

# Load data
load("../data/cleaned-data_jul23.rda")

# Discard low resolution interactions
int_good_res <- int_raw %>% 
  dplyr::filter(!str_detect(Prey, "\\*$")) %>% 
  dplyr::filter(!str_detect(Predator, "\\*$")) %>% 
  mutate(Prey = case_when(Prey == "Brachiopoda_spp" ~ "Brachiopoda", TRUE ~ Prey)) %>%
  dplyr::select(Prey, Predator, Reference, Reference_link) %>% 
  distinct(Prey, Predator, .keep_all = TRUE)

# Table S1. List of interactions
# knitr::kable(int_good_res %>% arrange(., (Predator)), booktabs = TRUE, format = "latex", col.names = c("Prey","Predator","Reference", "Link"), caption = "List of predator-prey (trophic) interactions for the food web of the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem. Alphabetically ordered by predator.") %>%
#    column_spec(1:3, width="2cm") %>%
#    column_spec(4, width="14cm") %>%
#    kableExtra::kable_styling(latex_options = c("hold_position","repeat_header"), font_size = 4) %>%
#    kableExtra::landscape()

pandoc.table(int_good_res %>% arrange(., (Predator)) %>% mutate(Reference_link = paste('\\tiny', Reference_link), Reference = paste('\\tiny', Reference), Prey = paste('\\tiny', Prey), Predator = paste('\\tiny', Predator)) %>% rename("Link"="Reference_link"), style="multiline", split.tables = Inf, justify = c("left","left","center","right"),
             caption = "List of predator-prey (trophic) interactions used to build the network for the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem (MPAN-BB).")

```

\newpage

```{r tableS2, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Load packages
# library(knitr)
# library(dplyr)
# library(kableExtra)
# library(stringr)

# Load data
load("../data/cleaned-data_jul23.rda")
load("../data/foodweb-data_jul23.rda")

# Tidy data
df_g <- igraph::as_data_frame(g, 'both')
sp_list <- df_g[["vertices"]] %>% 
  rename("TrophicSpecies" = "name") %>% 
  left_join(sp_raw) %>% 
  dplyr::select(TrophicSpecies, FunctionalGroup, Reference_pres, Reference_pres_link, Observation) %>% 
  rename("Species" = "TrophicSpecies", "Reference" = "Reference_pres") %>% 
  mutate(Observation = case_when(Observation %in% c("Registrado Psolus_sp; Martinez sugiere presencia de P. segregatus", "Registrado Psolus_sp; Martinez sugiere presencia de P. patagonicus", "Registrado Psolus_sp; Martinez sugiere presencia de P. antarcticus", "Lista de sp no publicada provista por Díaz de Astarloa", "Info from Sup Info (Teso_etal_2019)", "AMPNBB-II", "Abundance data available ?", "Esquema trófico", "Depredadores tope", "Copepoda", "Benthic. Not recorded", "Cirripedia") ~ "", 
                                 Species == "Campylaspis_spp" ~ "Aggregated: Campylaspis_frigida & Campylaspis_bacescui",
                                 Species == "Ophiactis_sp" ~ "Aggregated: Ophiactis_asperula & Ophiactis_spp",
                                 Species == "Brachiopoda_spp" ~ "Aggregated: Magellania_venosa, Liothyrella_uva & Terebratella_dorsata",
                                 Species == "Enteroctopodidae" ~ "Aggregated: Enteroctopus_megalocyathus & Muusoctopus_longibrachus",
                                 TRUE ~ Observation),
         Reference = case_when(Species == "Bacteria" ~ "Guinder_etal_2020",
                               Species == "Necromass" ~ "Riccialdelli_etal_2017", 
                               Species == "Syllidae" ~ "Schejter_etal_2016; Bremec_etal_2019",
                               Species == "Sabellidae" ~ "Schejter_etal_2016; Bremec_etal_2019; Spinelli_etal_2020",
                               Species == "Ophiactis_sp" ~ "Schejter_etal_2016; 2020a; Brogger_pers.com.",
                               Species == "Campylaspis_spp" ~ "Doti_etal_2020",
                               Species == "Brachiopoda_spp" ~ "Schejter_etal_2016", 
                               Species == "Salpa_thompsoni" ~ "Capitanio_pers.com.",
                               TRUE ~ Reference),
         FunctionalGroup = case_when(Species == "Campylaspis_spp" ~ "Cumacea",
                                     Species == "Brachiopoda_spp" ~ "Brachiopoda", TRUE ~ FunctionalGroup)) %>% 
  mutate(TaxLevel = case_when(str_detect(Species, "_") | str_detect(Species, "_spp") ~ "Species",
                              str_detect(Species, "_sp") ~ "Genus",
                              str_detect(Species, "dae$") ~ "Family",
                              Species %in% c("Detritus","Necromass") ~ "NA",
                              Species == "Foraminifera" ~ "Phylum",
                              Species == "Bacteria" ~ "Domain",
                              Species %in% c("Calcarea","Hexactinellida") ~ "Class",
                              Species %in% c("Halichondrida","Haplosclerida","Poecilosclerida","Tetractinellida") ~ "Order",
                              Species %in% c("Copepodite","Larvae_Fish","Larvae_Nauplii","Eggs_Fish") ~ "Ontogenetic stage",
                              TRUE ~ "Other")) %>%  
  dplyr::select(Species, TaxLevel, FunctionalGroup, Reference, Reference_pres_link, Observation) %>% 
  arrange(., (Species))

# Food web resolution
res_sp <- sum(sp_list$TaxLevel == "Species") / nrow(sp_list)

# Table 2. List of species
# knitr::kable(sp_list, format = "latex", booktabs = TRUE, linesep = "", longtable = TRUE, col.names = c("Species","Taxonomy","Group","Reference","Link","Observation"), caption = "List of species for the food web of the Marine Protected Areas Namuncurá - Banco Burdwood I and II ecosystem. Alphabetically ordered by species.") %>%
#   column_spec(1:3, width="2cm") %>%
#   column_spec(4, width="3cm") %>% 
#   column_spec(5:6, width="5cm") %>% 
#   kable_styling(latex_options = c("hold_position", "repeat_header"), font_size = 4) %>% 
#   landscape()

pandoc.table(sp_list %>% arrange(., (Species)) %>% mutate(Species = paste('\\tiny', Species), TaxLevel = paste('\\tiny', TaxLevel), FunctionalGroup = paste('\\tiny', FunctionalGroup), Reference = paste('\\tiny', Reference), Reference_pres_link = paste('\\tiny', Reference_pres_link), Observation = paste('\\tiny', Observation)) %>% rename("Taxonomy"="TaxLevel", "Group"="FunctionalGroup", "Link"="Reference_pres_link"), style="multiline", split.tables = Inf, justify = c("left","left","left","left","left","left"),
             caption = "List of species for the food web of the Marine Protected Areas Namuncurá - Banco Burdwood I and II ecosystem. Alphabetically ordered by species.")

```

\newpage

## Results

Table 3 presents an exhaustive list of the species-level properties for the food web of the MPA Namuncurá - Banco Burdwood I ecosystem.

```{r tableS3, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
require(kableExtra)

load("../results/summary_results_jul23.rda")

# Table 3. List of species
tbl_3 <- spp_total %>% 
  mutate(TopRole = case_when(TopRole == "hubcon" ~ "netcon", TRUE ~ TopRole)) %>% 
  dplyr::select(TrophicSpecies, FunctionalGroup, TotalDegree, NumPrey, NumPred, Between, Close, TL, meanTrophicSimil, TopRole) %>% 
  arrange(., desc(TotalDegree)) %>% 
  mutate(across(where(is.numeric), round, 4),
         FunctionalGroup = case_when(TrophicSpecies == "Brachiopoda_spp" ~ "Brachiopoda",
                                     TrophicSpecies == "Campylaspis_spp" ~ "Cumacea",
                                     TRUE ~ FunctionalGroup))

pandoc.table(tbl_3 %>% arrange(., (TotalDegree)) %>% rename("Species"="TrophicSpecies", "Group"="FunctionalGroup", "Degree"="TotalDegree", "TS"="meanTrophicSimil", "TR"="TopRole"), style="multiline", split.tables = Inf, justify = c("left","left","center","left","left","left","left","left","left","left"),
             caption = "Species-level properties for the food web of the Marine Protected Area Namuncurá - Banco Burdwood I ecosystem. Ordered by degree (total number of interactions). Between: betweenness; Close: closeness; TL: trophic level; TS: trophic similarity; TR: topological role (netcon = network connector, modcon = module connector, modhub = module hub, modspe = module specialist).")

```

\elandscape
\newpage

Figure 2 shows the species' topological roles in the food web graph. Species, represented by circles, are coloured by role: network connector (dark purple), module connector (light pink), module hub (light green), and module specialist (dark green). Network connector = 1 species; module connector = 41 species; module hub = 10 species; and module specialist = 328 species.

```{r toprole_graph, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 20, fig.height=20, fig.cap="Graph of the food web showing species coloured by topological role: network connector (dark purple), module connector (light pink), module hub (light green), and module specialist (dark green). Circle diameter is relative to the number of interactions."}

# Load packages
library(igraph)
library(multiweb)
library(grid)
library(dplyr)
library(ggplot2)

# Load data
load("../results/summary_results_jul23.rda")

# Topological roles
top.role.col <- hcl.colors(4, palette = "Purple-Green")
TRColor <- setNames(top.role.col, levels(as.factor(V(g_up)$TopRole)))
TRColor <- as.data.frame(TRColor) %>% 
  mutate(TopRole = c("hubcon", "modcon", "modhub", "modspe"))
#show_col(top.role.col)  # shows color & code
data_TRColor <- spp_total %>% 
  dplyr::select(TopRole, everything()) %>% 
  left_join(TRColor)
g_up_c <- g_up %>% 
  set_vertex_attr("TRColor", value = data_TRColor$TRColor)

# Number of species by Role
TR_table <- spp_total %>% 
  group_by(TopRole) %>% 
  summarise(n = n()) %>%
  mutate(freq = n / sum(n)) %>% 
  mutate(TopRole = case_when(TopRole == "hubcon" ~ "Network connector",
                             TopRole == "modcon" ~ "Module connector",
                             TopRole == "modhub" ~ "Module hub",
                             TopRole == "modspe" ~ "Module specialist"))

# Filter network connectors ('hubcon') & module connectors ('modcon'). Violet & pink respectively
sp_con <- spp_total %>% 
  filter(TopRole == "hubcon" | TopRole == "modcon")

# Move to a new page
grid.newpage()
# Arrange figure
par(cex.axis=1.5, cex.lab=1.5)
plot_troph_level(g_up_c, vertexLabel = F, modules = F, vertex.color = V(g_up_c)$TRColor, ylab = "Trophic level")

```

\newpage

## References